### 2018-07-03 paper/Version14/draft_v14 (Shannon Gallagher's conflicted copy 2018-06-26).tex

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
\usepackage{systeme}
%% tables
\usepackage{booktabs}
\usepackage{lscape}
\usepackage[table,xcdraw]{xcolor}
\usepackage{float}
\usepackage{url}

%% Make red text
\newcommand{\com}[1]{\textcolor{red}{ #1}}

%\linespread{2}

\usepackage[authoryear]{natbib}

\usepackage{tikz}
\tikzset{
  int/.style={circle, draw, fill=blue!20, minimum size=3em},
  init/.style={pin distance=1.2cm,pin edge={loop,thin,black}}
}
\usetikzlibrary{arrows,automata}

\pagenumbering{arabic}

\newcommand{\XX}{\ensuremath{25}} % number of methods

\newcommand{\xxsir}{\ensuremath{9} } % number of SIR methods
\newcommand{\wxxsir}{nine } % lower case number as word
\newcommand{\Wxxsir}{Nine } % capitalized word

\newcommand{\rr}{\ensuremath{\mathcal{R}_0}}

\setlength{\itemsep}{0pt}

\begin{document}

%%% draw a tree
\tikzset{
  treenode/.style = {shape=rectangle, rounded corners,
                     draw, align=center,
                     top color=white, bottom color=blue!20},
  root/.style     = {treenode, font=\Large, bottom color=red!30},
  env/.style      = {treenode, font=\ttfamily\normalsize},
  dummy/.style    = {circle,draw}
}

\title{\Wxxsir methods to estimate $\rr$ in the SIR model}
\author{ Shannon Gallagher$^{\dag}$, Andersen Chang$^{\ddag}$, and William F. Eddy$^{\dag}$ \\$\dag$ Department of Statistics and Data Science, Carnegie Mellon University\\ $\ddag$ Department of Statistics, Rice University}
\date{\today}
\maketitle

% \tableofcontents

\section{Introduction}\label{sec:intro}
What has been called ``arguably the most important quantity in the study of epidemics'' \citep{Heesterbeek2002},  $\mathcal{R}_0$, the reproduction number (by convention pronounced ``R-naught'') is an elusive quantity for epidemic modelers to estimate.  As defined by \citet{anderson1992}, $\rr$ is the ``the average number of secondary infections produced when one infected individual is introduced into a host population where everyone is susceptible.''  In some ways,$\rr$ summarizes an entire outbreak of a disease; it  is used to assess whether a disease epidemic will occur.  Additionally, it describes what percentage of the population needs to be vaccinated to avoid such an epidemic, roughly $1-\rr^{-1}$ \citep{anderson1992}).  Despite a clear definition of $\rr$, epidemiologists have struggled to create a standard  estimator for $\rr$  \citep{hethcote2000}.  A major issue in estimating $\rr$ is that the quantity is a \textit{property of the model}, meaning that the estimate $\hat{\rr}$ is dependent not only on the usual noise that comes with statistical modeling but also on a variety of assumptions on how researchers assume a disease is transmitted through a population \citep{diekmann2009}.

To elaborate on $\rr$ being a property of the model, we first need to introduce the concept of the ``SI-framework,'' which was pioneered by Kermack and McKendrick in the 1920s \citep{getz2006}.   Here, S and I are known as compartments where S stands for ``susceptible'' and I for ``infectious.'' Infectious disease models then specify how individuals move from the S compartment to the I compartment.  Often, ancillary compartments are added such as R, which stands for recovered or removed, or E, which stands for exposed but not yet infectious.  The estimate of $\rr$ from a SIR model will not be directly comparable to the estimate of $\rr$ from a SEIR model, for example.  The difficulties of estimating $\rr$ are summarized and expanded upon by \cite{li2011}.  \cite{driessche2017} overviews a number of methods used to estimate $\rr$ but looks at systems beyond the SIR model and so these methods may not be directly compared.   As a result of $\rr$ being a property of the model, we limit our review to methods for estimating $\rr$ for the SIR model introduced by \cite{Kermack700}.

Even when limiting $\rr$ estimates to those from SIR models, estimation is still difficult.  Problems include mathematical versus statistical methods (i.e. solving for $\rr$ versus estimating $\rr$), the number of observations used to estimate $\rr$ (i.e. time), boundary cases of infection and recovery rates, population size, initial SI ratio, and restrictive assumptions placed upon the noise.  These are all issues for producing point estimates, to say nothing of confidence or credible intervals (CI).

We review \wxxsir methods for estimating $\rr$.  We note that these methods are not exhaustive but are chosen due to their impact and use in epidemiology.  We find that these methods also help to highlight the problems mentioned above.

In addition to making a point estimate of $\rr$, we are particularly interested in the hypothesis  $H_0: \rr < 1$, as a value less than 1 indicates that a disease will not outbreak.  We also discuss techniques for estimating CIs for each of the different methods.  These CI estimates are created using the delta method, the block bootstrap, and the posterior distribution \citep{cao1999,wasserman2004}.

In the following sections, we present a series of simulations in which we examine each of our $\rr$ estimation methods.  Within these simulations, we vary the number of observations used to estimate $\rr$, infection and recovery rates, the total population size, initial SI ratio, and several assumptions about the noise.  After analyzing our simulations, we apply our methods to data from the USA H1N1 influenza pandemic.

Finally, we discuss our recommendations for estimating $\rr$ in the SIR model, and we comment on how these recommendations extend to more complex epidemiological models beyond the SIR model and on how many of the problems with estimating $\rr$  become exacerbated in such models.

The rest of this manuscript is organized as follows.  In Section \ref{sec:r0}, we briefly discuss the origin of $\rr$.  In Section \ref{sec:sir-intro}, we introduce the SIR model as described by \cite{Kermack700}.  In Section \ref{sec:methods}, we overview the \wxxsir methods for estimating $\rr$. In Section \ref{sec:ci}, we discuss how to estimate CIs for each of the \wxxsir methods.  In Sections \ref{sec:sim-res} - \ref{sec:sim-var-res}, we present the results of our simulations.  In Section \ref{sec:real-data}, we evaluate the methods on real data--from the USA H1N1 influenza pandemic of 2009.  Finally, in Section \ref{sec:discussion}, we provide recommendations of estimating $\rr$ for the SIR model and offer comments on how these conclusions extend to more complex models.

\section{Origins and Difficulties of Estimating $\rr$}
\label{sec:r0}

The origins of $\rr$ are tied to the survival function, which originates from the field of demography in the late 1800s.  The survival function describes how many female offspring a woman is expected to produce in her lifetime, literally the reproduction number \citep{dietz1993estimation}.  In demography, we have
\begin{align}\label{eq:surv}
\rr = \int_0^\infty p(a) \beta(a) da
\end{align}
where $p(a)$ denotes the probability of a woman surviving to age $a$ and $\beta(a)$ the rate of an individual of age $a$ giving birth to a  girl.  The concept of the reproduction number was later imported to the field of epidemiology by MacDonald and Smith \citep{dietz1993estimation}.  Analogously in epidemiology, $p(a)$ is the age of a disease, and $\beta(a)$ is the infection rate at time $a$.

Equation \ref{eq:surv} is the chronologically first, and in many ways, the most direct method to estimate $\rr$.  The premise of this whole paper is based upon the fact that estimating $\rr$ from the survival function in Equation \eqref{eq:surv} is a difficult task.  It is due to this difficulty that other estimates even exist and why we must treat $\rr$ as a property of the SIR model.

Other difficulties in estimating $\rr$ arise from model assumptions on how randomness enters the model and sensitivity of the estimate to changes in the model parameters.  For example, the SIR model is special in that two of its compartments, S and R are monotonic.  The number of susceptibles is non-increasing and the number of recovereds is non-decreasing.  Logically, models should enforce this monotonicity of the two compartments.  As a consequence, this restricts assumptions on how randomness enters the model.  For example, Gaussian error, a common assumption in statistics and machine learning, becomes highly suspect due to the assymmetrical nature of the noise imposed by the monotonic $S$ and $R$ compartments.  Without assuming Gaussian noise, it becomes much more difficult to guarantee properties such as unbiased and consistent estimates of $\rr$.

These difficulties in estimating $\rr$ have led a number of researchers to analyze the sensitivity of $\rr$ with respect to small perturbations in parameters like $\beta$ and $\gamma$ \citep{lash2003,epstein2007agent,capaldi2012}.  Even this is a difficult task. Since there is no analytic solution to the SIR model, estimating the set of derivatives with respect to $\beta$ and $\gamma$ must be done numerically, which can substantially increase computation time.  We should note that this sensitivity analysis is done before even adding any distributional assumptions on how randomness enters the model.

Below in Table \ref{tab:r0-real-ex}, we provide examples of estimates of $\rr$ for specific diseases from publications over the past three decades, spanning various locations over the world.  We display estimates for HIV, Zika (ZIK-V), Ebola (EVD), seasonal influenza, H1N1 influenza, and measles.  These estimates of $\rr$ are not limited to those of SIR model, although most results come from this framework.  Of these, Zika seems to have the largest estimated values of $\rr$, followed by Ebola, H1N1 influenza, seasonal influenza, and HIV.  We see that African locations seem to have larger estimates of $\rr$ compared to the rest of the world, regardless of the disease.  Some of the accompanying intervals (not necessarily CIs) are extremely large such as the case of HIV in Uganda (Interval: [0.12, 14.17]) whereas some are extremely small such as for Ebola in Guinea (Interval: [1.50, 1.52]).  We see that estimates for H1N1 influenza are generally higher than that of seasonal influenza, which is evidence that H1N1 should have been treated as a more critical concern than the typical seasonal influenza cycle.  The lowest possible estimate for $\rr$ reported is 0.12 for HIV in Uganda, and the largest is 17.00 for measles in England and Wales.  Within diseases, we see that estimates for $\rr$ seem to be similar, although the respective CIs typically do not intersect, suggesting that there is disparity in estimates of $\rr$.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{graphicx}
% \usepackage{lscape}
\begin{landscape}
\begin{table}
\centering
\begin{tabular}{@{}lllrrrl@{}}
\toprule
\textbf{Disease}    & \textbf{Location}                & \textbf{Year(s)}                 & \textbf{$\rr$} & \textbf{Lower} & \textbf{Upper} & \textbf{Source}                                                                                                                  \\ \midrule
HIV                 & Belgium                          & 2004-2010                     & 1.10          & 1.09           & 1.16           & \cite{coelho2011}   \\
HIV                 & Netherlands                      & 2004-2010                     & 1.11        & 1.10            & 1.18           & \cite{coelho2011}    \\
HIV                 & Portugal                         & 2004-2010                     & 1.08        & 1.06           & 1.15           & \cite{coelho2011}   \\
HIV                 & San Francisco, USA               & 1993                          & 5.00           & 2.00              & 8.00              & \cite{blower1994}    \\
HIV                 & Uganda                           & 2002-2008                     & 3.31        & 0.12           & 14.17          & \cite{nsubuga2014}    \\ \hline
ZIK-V                & French Polynesia                 & 2013                          & 2.75        & 2.53           & 2.98           & \cite{zhang2017}        \\
ZIK-V                & French Polynesia                 & 2013-2014                     & 3.70         & 2.60            & 4.80            & \cite{kucharski2016}  \\
ZIK-V                & Florida                          & 2016                          & 0.16        & 0.13           & 0.19           & \cite{dinh2016}       \\
ZIK-V                & Colombia                         & 2015-2016                     & 2.50         & 3.00              & 3.60            & \cite{nishiura2016}    \\
ZIK-V                & Barranquilla, Colombia           & 2015                          & 2.40         & 3.80            & 5.60            & \cite{towers2016}    \\
ZIK-V                & Colombia                         & 2015                          & 1.42        & 2.56           & 3.83           & \cite{majumder2016}  \\ \hline
EVD               & Western Africa                   & 2014                          & 1.78        & 1.60            & 2.00              & \cite{fisman2014}   \\
EVD & Western Africa                   & 2014                          & 1.60         & 1.40            & 1.80            & \cite{towers2014}  \\
EVD                 & Democratic Republic of the Congo & 1995                          & 2.70         & 1.90            & 2.80            & \cite{legrand2007}  \\
EVD                 & Uganda                           & 2000                          & 2.70         & 2.50            & 4.10            & \cite{legrand2007}  \\
EVD                 & Guinea                           & 2014                          & 1.51        & 1.50            & 1.52           & \cite{althaus2014}   \\
EVD                 & Sierra Leone                     & 2014                          & 2.53        & 2.41           & 2.67           & \cite{althaus2014}      \\
EVD                 & Liberia                          & 2014                          & 1.59        & 1.57           & 1.60            & \cite{althaus2014}    \\ \hline
Seasonal Influenza  & USA, France, Australia           & 1972-2002 & 1.30         & 1.20            & 1.40            & \cite{chowell2008}  \\
Seasonal Influenza  & French Territory                 & 1987-1995                     & 1.50         & 1.09           & 1.73           & \cite{bonabeau1998}  \\ \hline
H1N1 Influenza      & Thailand                         & 2009                          & 2.07        & 1.92           & 2.22           & \cite{desilva2009}     \\
  H1N1 Influenza     & Peru                             & 2009                          & 1.37        & 1.20            & 1.70            & \cite{desilva2009}        \\ 
  H1N1 Influenza & USA & 2009 & 1.50 & 1.30 & 1.70 &\cite{towers2009} \\ \hline
Measles      & Cirencester, England                 & 1947-1950                          & 13.50   & 13.00           & 14.00            & \cite{fraser2009}     \\ 
  Measles     & England and Wales                  & 1950-1968                          & 17.00      & 16.00          & 18.00         & \cite{anderson1992}     \\
  Measles     & Kansas, USA                  & 1918-1921                          & 5.50    & 5.00          & 6.00          & \cite{anderson1992}     \\
  Measles     & Ontario, Canada                  & 1912-1913                         & 11.50        & 11.00           & 12.00            & \cite{anderson1992}     \\
  Measles     & Willesden, England                  & 1912-1913                          & 11.00        & 12.00       & 1.60            & \cite{anderson1992}     \\
  Measles     & Ghana                 & 1960-1968                          & 14.50       & 14.00          & 15.00            & \cite{anderson1992}     \\
  Measles     & Eastern Nigeria                  & 1960-1968                          & 16.50        & 16.00          & 17.00           & \cite{anderson1992}     \\
\bottomrule
\end{tabular}
\caption{Estimates of $\rr$ from various researchers for the infectious diseases of HIV, (Zika) ZIK-V, Ebola (EVD), seasonal influenza, and H1N1 pandemic influenza ranging from the years 1972-2016 in various locations.  We also report the lower and upper bounds reported by the researchers.}
\label{tab:r0-real-ex}
\end{table}
\end{landscape}

\section{Kermack and McKendrick's SIR Model}
\label{sec:sir-intro}

The SIR model introduced by \cite{Kermack700} is a compartment model, where individuals move from susceptible, to infectious, and finally recovered states.  We make five essential assumptions,
\begin{enumerate}
\item The compartments are discrete and have no overlap.
\item The transition of objects into and out of compartments is described by a set of known equations, possibly dependent on unknown parameters.
\item The populations mix homogeneously.
\item The number of objects in each compartment at time $t=0$ is known.
  \item The law of mass action is followed.
  \end{enumerate}

The law of mass action is a property borrowed from chemistry which says that the mass of the product per unit time is proportional to the mass of the reactants \citep{lotka1920}.  In epidemiology, this means that the proportion of new infections per unit time is proportional to the current  number of susceptible \citep{anderson1992}.

In this particular  SIR model (displayed in Figure \ref{fig::sir}), the number of individuals ($N$) is constant. Adaptations of the SIR model can include birth and death rates, which may correspondingly change the derivation of $\rr$ (for further discussion, see Section \ref{sec:discussion}). The parameters  $\beta$ and $\gamma$ have epidemiological meaning.  Here, $\beta$ is the inverse of expected time to infection rate and $\gamma$ is the inverse of the expected time to recovery.  The movement of individuals from one compartment to another is represented through the ordinary differential equations below.  Both $\beta$ and $\gamma$ are constrained to be between $0$ and $1$.  For the remainder of this paper, we use $X$, $Y$, and $Z$ to denote the number of individuals in the $S$, $I$, and $R$ compartments, respectively, to avoid any confusion with $\rr$.
\begin{align}
\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Y}. \label{eq:sir}
\end{align}
In words, susceptible individuals become infected at a rate that is proportional to the percentage of infected individuals multiplied by $\beta$, the inverse of expected time to infection, and the number of susceptible individuals.  Infectious individuals recover at a rate of $\gamma$ multiplied by the number of infected individuals.

The SIR model is represented graphically in Figure \ref{fig::sir}. 
\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=3cm,auto,>=latex',every node/.append style={align=center}]
    \node [int,  fill = white!70!blue] (a)              {$X$};
    \node [int,  fill = white!70!blue]           (c) [right of=a] {$Y$};
    \node [int,  fill = white!70!blue] (e) [right of=c] {$Z$};
    \path[->, auto=false] (a) edge node {} (c)
                          (c) edge node {} (e) ;

\end{tikzpicture}
\caption{Depiction of a SIR model where $X=S, Y=I,$ and $Z=R$.  One can only get the disease once in this model.}\label{fig::sir}
\end{figure}
An outbreak occurs if the rate of change of infectious individuals is positive, $\frac{dY}{dt} > 0$, or equivalently,
\begin{align*}
  \frac{dY}{dt} &> 0 \\
  \implies -\frac{dX}{dt} - \frac{dZ}{dt} &> 0 \\
  \implies \frac{\beta X Y}{N}  - \gamma Y &> 0 ,\\
\implies  Y \left ( \beta \frac{X}{N} - \gamma \right ) & > 0\\
\implies   \frac{\beta}{\gamma} &> \frac{N}{X}.
\end{align*}
That is,  the rate of new infections is greater than the rate of recovery.  So as long as the number of initially susceptible individuals is large compared to the total population, $\frac{N}{X} \approx 1$, then an outbreak will occur if $\rr >1$,
\begin{align*}
  \rr \overset{def}{=} \frac{\beta}{\gamma}.
  \end{align*}
  To incorporate randomness into the model, we add noise, namely,
  \begin{align}\label{eq:sir-noise}
    X_{obs}(t) &= X(t) + \epsilon_{X,t}\\
    Y_{obs}(t) &=  N - X_{obs}(t) -Y_{obs}(t)  \nonumber\\
    Z_{obs}(t) &= Z(t) - \epsilon_{Z,t}. \nonumber
  \end{align}
We are assuming the observations are generated based on the discretized time (replacing $d$ with $\Delta$), deterministic ODEs presented in Equation \ref{eq:sir} with the addition of time ($t=0, 1, \dots, T$) and compartment dependent noise $\epsilon_{X,t}$ and $\epsilon_{Z,t}$.  Since, $N$, the total population, is constant, then $Y_{obs}$ is adjusted accordingly.

In summary, when we discuss estimators for $\rr$ for the SIR model, we mean to say we are forming an estimator of $\rr$ from the given set of data in Eq. \ref{eq:sir-noise},
\begin{align*}
  \textnormal{Data} &= \left \{\left (X_{obs}(t), Y_{obs}(t), Z_{obs}(t) \right ) : t=0, 1, \dots, T\right \}, \\
  \hat{\rr} &= m(\textnormal{Data}),
\end{align*}
where $m$ is a function of the data.

Good properties of an estimator may include (1) being a consistent estimator
\begin{align*}
  \hat{\rr} \overset{P}{\to} \rr \textnormal{ as } T\to \infty,
\end{align*}
where `$\overset{P}{\to}$' stands for convergence in probability \citep{wasserman2004} or (2) converging to a known probability distribution (convergence in distribution `$\overset{d}{\to}$' \citep{wasserman2004}),
\begin{align*}
\hat{\rr} \overset{d}{\to} F.  
\end{align*}
As we are interested in the hypothesis
\begin{align*}
  H_0:\;& \rr > 1 \\
  H_A:\;& \rr \le 1
\end{align*}
Then for some $\alpha \in [0,1]$, we would like to know for some $\hat{a}, \hat{b} >0$
\begin{align*}
P(\hat{a} \le \rr \le \hat{b}) \le 1 - \alpha.
\end{align*}
This property is particularly desirable, as we can estimate $\hat{a}$ and $\hat{b}$ from the distribution $F$.  At any rate, we would like to be able to estimate $\hat{a}$ and $\hat{b}$ for a given $\alpha$-level.  We could also make similar hypothesis tests such as that the $\rr$ for H1N1 influenza is greater than the reported $\rr$ for seasonal influenza.

\subsection{$\rr$ is a property of the model}\label{sec:prop}

We provide an example of how $\rr$ is a property of the model.  Consider the situation of data generated from an SEIR (XEYZ) model (described by \cite{cintronarias2009}), generated from the following ODEs,

\begin{align*}
  \frac{dX}{dt} &= - \frac{\beta XY}{N} \\
  \frac{dE}{dt} &= \frac{\beta XY}{N}  - \alpha E\\
  \frac{dY}{dt} &= \alpha E - \gamma Y \\
  \frac{dZ}{dt} &= \gamma Y,
\end{align*}
again, with a constant population of $N$ and with known initival values $(X(0), E(0), Y(0), Z(t))$.  The parameter $\alpha$ is the rate at which latent exposed individuals become infectious.  As $\alpha \to \infty$, the XEYZ model converges to the XYZ model.  The reproduction number for the XEYZ model, is the same as in the XYZ model, that is $\rr = \frac{\beta}{\gamma}$.

Consider the case when $N=10000, (\beta, \gamma) = (.06, .03)$, $(X(0), E(0), Y(0), Z(t)) = (9500, 0, 500, 0)$ for times $t=1, \dots, 365$.  Let the parameter $\alpha$ range between [0.005, 0.25] with step size $0.005$.  This generates a first data set
$$D_{\alpha, \textnormal{XEYZ}} = \left \{(X_{XEYZ}(t), E_{XEYZ}(t), Y_{XEYZ}(t), Z_{XEYZ}(t): t=1, \dots, 365\right \},$$
which is observed by researcher 1.  Now assume that researcher 2 misspecifies the model for this data as a XYZ model rather than a XEYZ model.  That is, she observes all exposed individuals as susceptible and generates a second data set
$$D_{\alpha, XYZ} = \left \{X_{XYZ}(t), Y_{XYZ}(t), Z_{XEYZ}(t) : t = 1, \dots, 365 \right \},$$
where $X_{XYZ}(t) = X_{XEYZ}(t) + E_{XEYZ}(t)$, $Y_{XYZ}(t) = Y_{XEYZ}(t)$, and  $Z_{XEYZ}(t) = Z_{XYZ}(t)$. This situation is illustrated in Figure \ref{fig:sir-vs-seir} with $\alpha = 0.01$.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{images/seir-sir-data.pdf}
  \caption{Top: XEYZ generated data over the course of a year.  Bottom: data tranformed from the XEYZ model to the XYZ model by combining all X and E compartments.}\label{fig:sir-vs-seir}
  \end{figure}

Both researchers want to estimate $\rr$ and so fit models to their data sets.  The details of how the model are not so important at the moment (but will be in the later sections), but to fairly compare assume both researchers use least squares to fit the models.  That is researcher 1 estimates $\rr$ with
\begin{align*}
  (\hat{\beta}_{XEYZ}, \hat{\gamma}_{XEYZ}, \hat{\alpha}_{XEYZ} )&=\arg \min_{\beta, \gamma, \alpha} \sum_{t} \left [ \left (X_{XEYZ}(t) - X(t)\right )^2  \right . \\
                                                                 &+\left ( E_{XEYZ}(t) - E(t) \right )^2  \\
  &\left . + \left ( Z_{XEYZ}(t) - Z(t) \right )^2\right ]
\end{align*}
Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
\begin{align*}
  \hat{\rr}^{XEYZ}= \frac{\hat{\beta}_{XEYZ}}{\hat{\gamma}_{XEYZ}}.
\end{align*}
Researcher 2 estimates $\rr$ with 
\begin{align*}
(\hat{\beta}_{XYZ}, \hat{\gamma}_{XYZ} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{XYZ}(t) - X(t)\right )^2 + \left ( Z_{XYZ}(t) - Z(t) \right )^2 \right ].
\end{align*}
Then the least squares (LS) estimate for $\rr^{XYZ}$ is given by Equation \ref{eq:sirls},
\begin{align*}
  \hat{\rr}^{XYZ}= \frac{\hat{\beta}_{XYZ}}{\hat{\gamma}_{XYZ}}.
\end{align*}

We plot the two estimates of $\rr$ as a function of $\alpha$ in Figure \ref{fig:r0-alpha}.  We should note we also estimated the standard errors of the estimates of $\rr$, but they do not show up on the figure because they are of an order of magnitude of $10^{-4}$.  Unsurprisingly, under the SEIR model speficiation, $\hat{\rr}=2$ because there is no noise in the data, and the data is generated from an SEIR model.  However, the SIR model for small values of $\alpha$ can result in double the size of $\rr$!

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{images/r0-est.pdf}
  \caption{Estimates of $\rr$ from both the XEYZ and XYZ model specification for different values of $\alpha$ using Least Squares with $\beta=0.06$, $\gamma=0.03$, and $(X(0), E(0), Y(0)) = (9500, 500, 0)$.}\label{fig:r0-alpha}
  \end{figure}

For those readers not convinced by our small example, we offer this mathematical intuition of the next generation model (NGM) described by \cite{diekmann1990} and \cite{diekmann2009}.

 The NGM is a generalization of any compartment model at the infection-free steady state. This model solves the problem of having an expression for $\rr$ in terms of the epidemiological parameters for a general compartment model.  Originally, introduced by \citep{diekmann1990}, \cite{diekmann2009} posit a recipe to find $\rr$ for a wide class of compartment models, including SIR and SEIR. First, define the infected subsystem as ``the equations of the ODE system that describe the production of new infections and changes in state among infected individuals.''  Let $x = (C_1, C_2, \dots, C_m)$ where $C_i$ are the different compartments of the infected subsystem.  The steps to find $\rr$ are as follows:

\begin{enumerate}
\item ``Linearize the infected subsystem of nonlinear ODEs about the infection-free steady state''
\item Decompose the linearized infected subsystem into $(T + \Sigma )x$ where $T$ is a $m\times m$ matrix of transmissions and $\Sigma$ is a $m \times m$ matrix of transitions.
\item $\rr$ is the spectral radius (i.e. the dominant eigenvalue) of $-T \Sigma^{-1}$.  
\end{enumerate}

Here, $T_{ij}$ is the rate of transmission of \textit{newly} infected individuals in state $i$ created by individuals in state $j$.  $\Sigma_{ij}$ is the transition rate of individuals into compartment $i$ from compartment $j$.

This method has the assumptions of a basic compartment model: homogeneous mixing among populations and the law of mass action, i.e., that the number of infected individuals are proportional to the number of infected individuals at the previous step.   Additionally, this method is advantageous in that the matrix $-\Sigma^{-1}$ has intuitive meaning.  The $ij$th entry of $- \Sigma^{-1}$ is the ``expected time that an individual who presently has state $j$ will spend in infected state $i$.''

The NGM shows that $\rr$ is the spectral radius of a matrix composed from the transmission of disease from diseased individuals to susceptible individuals and transition of diseased individuals from among the compartments and is indeed a property of the model.

\section{Overview of \wxxsir methods to estimate $\rr$ for the SIR model}
\label{sec:methods}

In Section \ref{sec:prop} we described how $\rr$ is a property of the model.  To fairly compare estimation methods to one another, we restrict our methods to estimate $\rr$ to the SIR model.  Also in Section \ref{sec:prop}, we examined simulated data that was generated from a SEIR model without noise.  However, we know that data is very noisy.    In this section, we describe the \wxxsir methods to estimate $\rr$ for the SIR model in the presence of noise in detail, including advantages and disadvantages of each.

\subsection{Least Squares ($\beta$, $\gamma$) (LS)}\label{least-squares-beta-gamma}
Our first approach to estimate $\rr$ in the SIR model is to minimize the joint mean square error for the data collected at each time point and use the plug-in estimator found in Equation \ref{eq:sirls}.  In particular, we find

\begin{align*}
(\hat{\beta}, \hat{\gamma} )&=\arg \min_{\beta, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
\end{align*}
Then the least squares (LS) estimate for $\hat{\beta}$ and $\hat{\gamma}$ is then combined to give an estimate of $\hat{\rr}$ in Equation \ref{eq:sirls},
\begin{align}\label{eq:sirls}
  \hat{\rr}= \frac{\hat{\beta}}{\hat{\gamma}}.
\end{align}

The estimates from this method are often found via a grid search of $\beta$ and $\gamma$ values to minimize the squared error.  The estimates can also be found using more sophisticated optimization algorithms. Since we typically cannot explicitly write down the partial derivatives of $S$, $I$, and $R$ with respect to $\beta$ and $\gamma$, we can only use minimization algorithms which do not rely on an explicit gradient such as Nelder-Mead simplex minimization \citep{nelder-mead1965}.

A reasonable question one may ask is why use the $L_2$-norm as opposed to $L_1$ or some other similarity score.  A possible answer is that if we were to assume Gaussian noise, then the $L_2$-norm would be equivalent to the maximum likelihood estimation.  Another answer is that $L_2$ is a continuously differentiable function and hence is easier to compute, which allows for sensitivity analysis to  be conducted more easily.  Finally, LS can be done without writing down explicit assumptions about the noise within the model.  However, we cannot guarantee properties of consistency or convergence to a known distribution without more explicit assumptions.

On the other hand, least squares is easy to implement using software such as \texttt{R}'s \texttt{optim()} function, and as we see in Section \ref{sec:results} produces results comparable methods to other methods.  Least squares and similar variants have been used to estimate $\rr$ in \cite{majumder2016}.

\subsection{Reparametrized Least Squares ($\rr$, $\gamma$) (ReLS)}\label{reparametrized-least-squares-rux5f0-gamma}

In the previous method (Section \ref{least-squares-beta-gamma}), we estimated $\beta$ and $\gamma$ and then estimated $\rr$.  However, it is possible to directly estimate $\rr$ if we reparametrize the ODEs in Equation \eqref{eq:sir} directly with \(\rr\) and \(\gamma\), using the relation $\rr = \frac{\beta}{\gamma}$; this yields

\begin{align*}
  \left \{
  \begin{array}{cl}
    \frac{dX}{dt} &= - \rr \gamma Y \frac{X}{N}\vspace{.5em}\\
    \frac{dY}{dt} &=  \rr \gamma Y \frac{X}{N}  - \gamma Y\vspace{.5em} \\
    \frac{dZ}{dt} &=  - \gamma Y 
  \end{array}
  \right . .
  \end{align*}
We find
\begin{align}
(\hat{\rr}, \hat{\gamma} ) &= \text{argmin}_{\rr, \gamma} \sum_{t} \left [ \left (X_{obs}(t) - X(t)\right )^2 + \left ( Z_{obs}(t) - Z(t) \right )^2 \right ]
\end{align}
We use the $\hat{\rr}$ directly from the above estimation problem, which again can be found with a grid search or another optimization process.

We can also write $\hat{\rr}$ as a function of $\gamma^*$,
\begin{align*}
\hat{\rr}(\gamma^*)&= \text{argmin}_{\rr} \sum_{t} \left [ \left (X_{obs}(t) - X(t; \gamma^*)\right )^2 + \left ( Z_{obs}(t) - Z(t; \gamma^*) \right )^2 \right ].
  \end{align*}

It may be surprising that this method leads to different results than simply using least squares.  We also see that this method further results in different standard errors than in LS, which we discuss further in Section \ref{sec:results}.

Like LS, while we have no explicit parametric assumptions on the noise, we also are uncertain of any theoretical properties of our resulting estimator.  Likewise, the same difficulties in sensitivity analysis arise as in LS.

\subsection{Linear Model Approximation (LMA)}\label{linear-model-approximation-degree-10}

The SIR ODEs in Eq. \ref{eq:sir} have no known closed form solution, and so we use numerical integration to obtain approximate solutions.  In addition to this, data collected from real diseases are typically very noisy to begin with.  \cite{chang2017} showed that the SIR model may be well approximated by a linear model.  We use this method to estimate $\rr$.

Specifically, we estimate two polynomials in \(t\) with degree $K$  to \(X_{obs}\)
and \(Z_{obs}\) using least squares to find the coefficients $\{(\hat{x}_k,
\hat{z}_k)\}_{k=1, \dots, K}$,
\begin{align*}
\hat{X}(t) &= \sum_{k=0}^K \hat{x}_k t^k\\
{\hat{Z}}(t) &= \sum_{k=0}^K \hat{z}_k t^k
\end{align*}
Then, we estimate the derivatives as
\begin{align*}
\hat{X}^\prime(t) &= \sum_{k=1}^K k \hat{x}_k t^{k-1}\\
\hat{Z}^\prime(t) &= \sum_{k=0}^K k \hat{z}_k t^{k-1}
\end{align*}
An estimator for $\rr$ is derived from the ODEs in Equation \eqref{eq:sir},
\begin{align}
  - \frac{X^\prime}{Z^\prime}&= \rr \frac{X}{N} \nonumber\\
  \rr &=       -\frac{X^\prime}{
        Z^\prime} \cdot \frac{N}{X} \nonumber\\
  \hat{\rr} &= -\frac{\hat{X}^\prime(0)}{ \hat{Z}^\prime(0)} \cdot \frac{N}{\hat{X}(0)}. \nonumber
  \end{align}
  Here, $K$ is arbitrary and should be selected using some criterion such as AIC (REF).  Besides optionally deciding on the degree of polynomials to fit, this model is simple to implement and gives comparable results to using least squares with the SIR model.  The time $t=0$ is used to best capture the initial outbreak.

  An advantage of using this estimation method is that it is simple to implement using any linear modelling software.  On the other hand, we assume $X(t) = \hat{X}(t)$ and so if the data truly follows the SIR model, then we will likely have a biased result for $\rr$.  Using this method, we are able to check how sensitive $\rr$ is to the degree of the fitted polynomial.

\subsection{Linear Model Approximation, All Time Points (LMAT)}\label{linear-model-approximation-all-time-points-degree-10}

The above formulation (Section \ref{linear-model-approximation-degree-10}) of a linear model approximation only uses the estimate at time $t=0$ to estimate $\rr$.  We can instead, use all time points $T$ available to estimate $\rr$.  We fit a polynomial in \(t\) with degree \(K\) to \(X_{obs}\)
and \(Z_{obs}\) as above, with a slight modification in how we estimate
\(\rr\),
\begin{align*}
  \hat{\rr} &= \frac{1}{T} \sum_t \frac{-\hat{X}^\prime(t)}{\hat{Z}^\prime(t)} \cdot \frac{N}{X(0)} 
\end{align*}
The intuition is that $\frac{-X^\prime(t)}{Z^\prime(t)}$ is constant in $t$, but due to our approximations with the linear model, this is no longer the case.  Here, we average over the different possible values of $\rr$, estimated at different times.  An advantage to this approach is that we have a more robust estimate of $\rr$ than just using one time point.  As in LMA, we can examine how sensitive $\rr$ is to the order of the polynomial $K$.

\subsection{Incidence to Prevalence Ratio (IPR)}\label{incidence-to-prevalence-ratio}
The incidence to prevalence ratio (IPR), described by \cite{Nishiura2009}, is another intuitive method to estimate $\rr$.  It incorporates some of the most basic epidemiological quantities: incidence and prevalence.

In terms of data from the SIR model, the incidence $J(t) \approx -(X(t+1) - X(t))$, and the IPR$(t) = \frac{J(t)}{Y(t)}$.  This method assumes that we have some prior knowledge about $\gamma$, the recovery rate.  Thus we use as our estimate,
\begin{align*}
\hat{\rr} &= \textnormal{IPR}(t) \cdot \frac{1}{\gamma}
\end{align*}

Here we assume that the time step is small enough to approximate the incidence.  The advantage of this method is that incidence data is generally readily available as is prevalence data for certain diseases such as HIV.  However, as one is required to have prior knowledge about $\gamma$, it may be easier to directly estimate $\rr$ with one of the many other methods described that does not require  prior knowledge.  Again, we are using only one time point to estimate $\rr$.  This model is agnostic to assumptions on the noise, but again, it is unclear if this is an unbiased or consistent estimator of $\rr$.

\subsection{Smoothed Incidence to Prevalence Ratio (SIPR)}
We use the same method as above, IPR, but first estimate splines with $K$ degrees of freedom, where $K$ is small, $\hat{X}(t)$ and $\hat{Y}(t)$, to fit to $X_{obs}$ and $Y_{obs}$, respectively.  Then  $J(t) \approx -(\hat{X}(t+1) - \hat{X}(t))$, and the $\hat{\textnormal{IPR}}(t) = \frac{J(t)}{\hat{Y}(t)}$.  Then
\begin{align*}
\rr &= \hat{\textnormal{IPR}}(t) \cdot \frac{1}{\gamma}
\end{align*}
Splines are used because they typically can be used to make good approximations of epidemic curves (CITE).  However, the researcher could use their method of choice to estimate $\hat{X}(t)$ and $\hat{Y}(t)$.  An advantage of this method is that it creates a less variable estimate than estimating IPR using only one point.  It has the same disadvantages as the usual IPR ratio in that it requires knowledge about $\gamma$.

%%%%%%%%%%%%%%%%%%%%%
\subsection{Log-Linear (LL)}
\cite{harko2014exact} were able to reduce the SIR model to one ODE.  From this, we can derive the following,
\begin{align}
  X(t) &=  X(0) e^{\frac{-\beta}{\gamma}\frac{Z(t)}{N}} \nonumber\\
  \log \frac{X(t)}{X(0)} &=  \frac{-\beta }{\gamma}\frac{Z(t)}{N} \nonumber\\
  \log \frac{X(t)}{X(0)} &=  -\rr \frac{Z(t)}{N}. \label{eq:harko_lin}
\end{align}

We can regress the left hand side in Eq. \ref{eq:harko_lin} on $Z(t)$ (using least squares), the number of recovered individuals, with $\rr$ as the coefficient of $Z(t)/N$ and with an intercept term of $\beta_0=0$ to obtain an estimate of $\rr$,
\begin{align*}
  \hat{\rr} = -\frac{\sum_{t=0}^T \log \frac{ X(t)}{X(0)}}{\sum_{t=0}^T\frac{Z(t)}{N}}.
\end{align*}
This method says that for a one percent increase in number of recovered individuals, we expect the ratio of the number of previous susceptibles to number of new susceptibles to increase by $e^{\rr}$,
\begin{align*}
  \log \left ( X(t+1)/ X(0) \right ) - \log \left ( X(t)/X(0) \right ) &= - .01\rr\\
  \log \left ( X(t+1) \right ) - \log \left ( X(t) \right )  &=- .01\rr\\
  \log \left ( X(t) / X(t+1) \right ) &= .01\rr\\
  \frac{X(t)}{X(t+1)}  &\approx e^{\rr}.
\end{align*}

The advantages of this method are many.  One, the estimate of $\rr$ is highly interpretable.  Two, an assumption of Gaussian noise in Equation \ref{eq:harko_lin} is not as strong as an assumption in the other methods to estimate $\rr$ because we are looking at a log transformation of the percent of suceptibles instead of raw variables.  If the assumption holds, we can show that the resulting estimate of $\rr$ is unbiased and consistent.  Finally,  this method is easily implemented using any linear modeling software.

\subsection{Markov chain estimation (MC)}
A natural approach to epidemic modelling is that of Markov chains (MC), since it is assumed an individual's next state is only dependent on its current state and the current states of other individuals.  Much work has been done over the years in this specific field including asymptotic behavior, continuous time MC, confidence intervals, and more \citep{jacquez1991,gani1995,daley2001epidemic}.  We present one simple instantiation of the model, the discrete time case, which traces its origin back to the Reed-Frost model \citep{abbey1952}.

In this method, the number of susceptibles at the next step, $X_{obs}(t+1)$, has a Binomial distribution based on the contacts with the current number of infectious, $Y(t)$ and the current number of susceptibles.  That is $X_{obs}(t+1) \sim \text{Binomial}\left(X_{obs}(t), \alpha^{Y_{obs}(t)}\right)$, where $\alpha$ is the probability of avoiding infection from an infective.  One can write out the likelihood for $\alpha$ in this model, namely,
\begin{align*}
\mathcal{L}\left ( \alpha ; \textnormal{data}\right ) \propto \prod_{t=0}^{T-1}\left ( \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t+1)} \left (1- \alpha^{Y_{obs}(t)} \right )^{X_{obs}(t) - X_{obs}(t+1)}.
\end{align*}
Maximizing the likelihood yields an estimate for $\alpha$,
\begin{align*}
\hat{\alpha} = \arg \max_{\alpha} \mathcal{L}\left(\alpha; \textnormal{data} \right ).
  \end{align*}
 \cite{barbour2004} report that the estimated reproduction number is thus,
\begin{align}\label{eq:r0-mc}
\hat{\rr} &= \log \left ( \frac{1}{1-\hat{\alpha}}\right ).
\end{align}

This method typically allows for more than just the reproduction number to be estimated.  Through recursion, one can estimate the probability of having a given number of susceptibles and infected at each time step, and hence the entire probability distribution may be known.

The advantages of this model are its simplicity of interpretation and ability to generate an estimate for the probability distrubtion for $(X(t), Y(t))$.  We have explicitly stated how randomness enters the model so it is up to the researcher to check whether this assumption is a good fit.  Using this method, it is easy to see how sensitive $\hat{\rr}$ is to changes in $\hat{\alpha}$ as $\hat{\rr}$ is a function of $\alpha$.

\subsection{Sequential Bayes (SB)}\label{sec:seqbayes}

Described by \cite{bettencourt2008} and summarized in \cite{obadia2012r0}, the sequential Bayes method is a Bayesian approach to an approximation of the classic SIR model.  The approximated SIR model assumes that the incidence at $t+1$, $J(t+1)$ has a Poisson distribution, with $\gamma$ as the  average inverse of the infectious period. In order to estimate $\rr$, we must have some idea about $\gamma$,
\begin{align*}
J(t+1) | \rr  \sim \textnormal{Poisson}( J(t) \exp \left \{  \gamma (\rr-1)\right \})
\end{align*}
Then, the posterior distribution of $\rr$ given the previous time period's incidence is
\begin{align*}
  P(\rr | J_0, \dots, J_{t+1}) = \frac{P(J_{t+1} | \rr, J_0, \dots, J_t)P(\rr| J_0, \dots, J_t)}{P(J_0, \dots, J_{t+1})}.
\end{align*}
This method is sequential in that the prior distribution for $\rr$ comes from the previous time period.  The initial prior for $\rr$ is assumed to be flat.  This method results in a posterior distribution from which credible intervals may be obtained.  This method assumes, initial growth in incidence to be exponential, and homogeneous mixing of populations as with any compartment model.  The advantages of this method are that of the ability to obtain an entire posterior distribution, whereas many other methods are difficult to even find an estimate of the variance.  Disadvantages include strict assumptions about the distributions and computational time required to estimate the relevant parameters.

Although, the Poisson distribution shown above is often used to estimate $\rr$, we have found that this model does not perform well in practice (see Section \ref{sec:results}).  Thus, we would recommend using a more sophisitcated Bayesian model such as the one in \cite{bauer2017}.

\section{Confidence and Credible Intervals}
\label{sec:ci}

Estimating $\rr$ is difficult and estimating $V[\rr]$, the variance and confidence or credible intervals (CI) is even more so.  We describe general methods which may be applicable to estimate the variance.  In particular, we describe three methods 1) the delta method, 2) the block bootstrap, and 3) posterior distributions.

\subsection{Delta Method}\label{delta-method}

When the method estimates \(\beta\) and \(\gamma\) instead of \(\rr\) directly, we use the delta method approximation to estimate the
variance of \(\rr\).   From \cite{wasserman2004}, if
\begin{align*}
  \sqrt{T} \left ( \hat{\mathbf{\theta}}_o - \mathbf{\theta} \right ) \overset{d}{\to} N\left ( \mathbf{0}, \Sigma \right  )
\end{align*}
and $h$ is a continuously differentiable function then
\begin{align*}
  \sqrt{T} \left (h(\hat{\theta}_n) - h(\theta) \right ) \overset{d}{\to} N \left (0,  \nabla^T h(\theta)\Sigma \nabla h(\theta)\right ),
\end{align*}
where $\nabla h$ is the gradient of $h$.

For the method LS, we know $\rr = h(\beta, \gamma) = \frac{\beta}{\gamma}$. Then, if $E\left [\frac{\hat{\beta}}{\hat{\gamma}}\right] \to \frac{\beta}{\gamma}$, then  $(\nabla h = (\frac{1}{\gamma},  -\frac{\beta}{\gamma^2})^T)$ and
\begin{align*}
  V[\rr] &\to \nabla h^T(\beta, \gamma) \Sigma_{\beta, \gamma} \nabla h(\beta, \gamma) \\
  &= \frac{\Sigma_{\beta, \beta}}{\gamma^2} - \frac{\Sigma_{\beta, \gamma}(1 + \beta)}{\gamma^3} + \frac{\beta \Sigma_{\gamma, \gamma}}{\gamma^4}
\end{align*}
where $\Sigma_{\beta, \gamma}$ is covariance matrix of \(\beta\) and \(\gamma\). .  We can use plug-in estimators for $\hat{\beta}$, $\hat{\gamma}$, and $\hat{\Sigma}$ where $\hat{\Sigma}$ is taken to the inverse of the Hessian from the least squares optimzation of $\beta$ and $\gamma$.

For ReLS, we use the estimate of $\Sigma_{\rr, \gamma}$ taken from the optmization of $\rr$ and $\gamma$ as to find our estimate for $\hat{\Sigma}_{\rr}$.

This estimate assumes that the distribution of $\beta$ and $\gamma$ are asymptotically normal.  Here, we use the relationship of $\rr = \frac{\beta}{\gamma}$ in order to estimate the variance, which is specific to the SIR model.  This method may be extended to other frameworks such as the SEIR model.  The advantages of the delta method are that since the Next Generation Matrix \citep{diekmann2009} gives a recipe for deriving $\rr$ from the deterministic ODEs, then it is theoretically possible to use the delta method to derive an expression for the variance of $\rr$ for any compartment model.

We can also use the Delta method to derive variance estimates of LMA, LMAT, LL, and MC as for all these estimates first estimate parameters $\theta$ since $\rr= h(\theta)$

\subsection{Block Bootstrap}

The block bootstrap is a variant of the bootstrap in which the $n$ observations are dependent on one another.  In contrast to the original bootstrap, the block bootstrap partitions consectutive observations into $k$ blocks with block-length $b$ ($n=kb$).  These are non-overlapping partitions or blocks, although there are variants for which blocks can overlap.  For iteration $\ell$, one samples $k$ blocks uniformly with replacement and estimates $\hat{\rr}_b$, repeating the sampling and estimation for a total of $L$ estimates of $\hat{\rr}$.  Then
\begin{align*}
  V\left [ \hat{\rr} \right ] &\approx V\left [\hat{\rr}_\ell \right ], \; \ell = 1, \dots, L
\end{align*}
More about the asympotic distributions of the blockwise-bootstrap can be found in \cite{cao1999}, along with descriptions of variants of the block bootstrap.

The asympotic properties hold when the stochastic process is $m$-dependent, that is when observations $t, t+1, \dots$ and $s, s+1, \dots$ are independent whenever $s+m < t$ and under some smoothness conditions on the statistic used to estimate $\rr$.

A choice one must make in the block bootstrap is the block size $b$.  One simple recommendation is to take $b= Cn^{1/3}$, which is derived from the asymptotic mean square error of the method.  \cite{hall1995} describe how to calculate $C$, which is context dependent.

\subsection{Posterior Distribution}
In this setting, we assume $\rr$ is a random variable and the data is fixed.  Following the notation of \cite{wasserman2004}, then the posterior distribution $f(\rr | \textnormal{data})$ is proportional to the likelihood of observing the data given $\rr$, $\mathcal{L}(\textnormal{data}|\rr)$, multiplied by our prior on $\rr$, $f(\rr)$. 
\begin{align*}
f(\rr | \textnormal{data} ) \propto \mathcal{L}(\textnormal{data} | \rr) f(\rr)
\end{align*}
 Typically, the posterior is either estimated using conjugate distributions for the likelihood and prior or through Markov Chain Monte Carlo (MCMC) to simulate a posterior distribution $\hat{F}$ such that
\begin{align*}
\hat{F} \overset{d}{\to} f(\rr| \textnormal{data}).
\end{align*}
Then commonly, $\hat{\rr} = \hat{F}_{.5}$, the median of $\hat{F}$.  Correspondingly, the 95\% credible interval is $\left[\hat{F}_{.025}, \hat{F}_{.975} \right ]$. Advantages of this method include having an entire distribution instead of a point estimate.  A disadvantage is that using conjugate priors may not be justified for the given data and MCMC may be computationally intractable.  One should note that these 95\% credible intervals are not the same as confidence intervals since the ``true'' value of $\rr$ may not be contained in the credible intervals 95\% of the time.  We only use the posterior distribution to form a CI for the SB method, as it is the only Bayesian method presented here.

%%%%%%%%%%%%%%%

\section{Methods}\label{sec:sim-res}

We compare the performance of the SIR models on simulated data from both the SIR as well as other models. In all of these simulations, we first generate data from the models under known conditions, then we add noise drawn from known distributions to each time point,
\begin{align}\label{eq:sim-models}
  X_{model}(t) &= f(t) + \epsilon_{X,t} \\
  Y_{model}(t) &= g(t) + \epsilon_{Y,t} \nonumber\\
  Z_{model}(t) &= N - X_{model}(t) - Y_{model}(t)\nonumber 
\end{align}
We adjust the recovered compartment $Z$ so the total population is constant. 
We simulate with two different types of error distributions, Gaussian (Norm) and autoregressive (AR),

\noindent \textbf{Gaussian (Norm) ($X_{G}$, $Y_{G}$, $Z_{G}$)}
\begin{align*}
  f(t) &= X(t) \\
  g(t) &= Y(t) \\
  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2)
\end{align*}
\textbf{Autoregressive (AR) ($X_{AR}$, $Y_{AR}$, $Z_{AR}$)}
\begin{align*}
  f(t) &= \rho X_{obs}(t-1) \\
  g(t) &= \rho Y_{obs}(t-1) \\
  \epsilon_{X,t} &\sim N(0, \sigma_X^2) \\
  \epsilon_{Y,t} &\sim N(0, \sigma_Y^2).\\
\end{align*}
For each of the Gaussian and AR datasets, we form an additional dataset which enforces the monotonicity of X and Z via the Pool Adjacent Violators Algorithm (PAVA), described by \cite{friedman1984}).  Here, we adjust $Y$ so the constant population is maintained.

\noindent \textbf{Gaussian-PAVA (Norm-M) ($X_{GP}$, $Y_{GP}$, $Z_{GP}$)}
\begin{align*}
 X_{GP}(t) &= \textnormal{PAVA}(X_G(t)) \\
  Y_{GP}(t) &= N - X_{GP}(t) - Z_{GP}(t) \\
  Z_{GP}(t) &= \textnormal{PAVA}(Z_G(t)) \\
\end{align*}
\textbf{Autoregressive-PAVA (AR-M) ($X_{ARP}$, $Y_{ARP}$, $Z_{ARP}$)}
\begin{align*}
  X_{ARP}(t) &= \textnormal{PAVA}(X_{AR}(t)) \\
  Y_{ARP}(t) &= N - X_{ARP}(t) - Z_{ARP}(t) \\
  Z_{ARP}(t) &= \textnormal{PAVA}(Z_{AR}(t)) 
\end{align*}
This gives us a total of four error structures for each set of conditions: (1) Gaussian (Norm), (2) Gaussian monotone (Norm-M),  (3) autoregressive (AR), and (4) autoregressive monotone (AR-M) for a given set of parameters $\left (\beta, \gamma, O, X(0), Y(0), Z(0), \sigma_X, \sigma_Y, N \right )$, the infection rate; recovery rate; total number of time points; initial conditions for number of susceptibles, infectious, and recovered individuals, the standard deviation for both the susceptible and infectious compartments, and the total population, respectively.

\subsection{SIR Data}

We first simulate a ``baseline'' data set from the SIR model under the following conditions:

\textbf{Baseline Simulated Data}
\begin{center}

	$\beta$ = 0.06, $\gamma$ = 0.03 ($\rr$ = 2)

	Number of time steps $T$: 365

	Starting X size ($X(0)$): 99950

	Starting Y size ($Y(0)$): 50

	Starting Z size ($Z(0)$): 0

	$\sigma_X$: 100

	$\sigma_Y$: 5

	Total population $N$: 100,000

        Error: Gaussian.

      \end{center}
      The baseline data set is generated from the deterministic SIR curves in Figure \ref{fig:baseline-data} and then noise is added to two of the compartments, and the third is adjusted to maintain a constant population value.  In this scenario, we see that the infection begins slowly but over the course of a year infects over 75\% of the population.
      \begin{figure}
        \centering
        \includegraphics[width=.8\textwidth]{images/baseline_sir.pdf}
        \caption{SIR (XYZ) curves for the deterministic model with $N=100,000$, $(\beta, \gamma) = (0.06, 0.03)$, $(X(0), Y(0), Z(0))= (99,950, 50,0)$  The baseline data is generated by using this model and adding noise to the $X$ and $Y$ compartments and adjusting $Z$ at each time step to maintain a constant population.}\label{fig:baseline-data}
        \end{figure}
      This baseline data set is chosen due to its moderate $\rr$ size of 2, which indicates an outbreak of an epidemic similar to HIV, ZIK-V, EVD, and H1N1 pandemic influenza (see Table \ref{tab:r0-real-ex}).  The infection rate and recovery rate ($\beta, \gamma$) are relatively small.  The duration of the disease, under the true model, spans nearly a whole year (see Figure \ref{fig:baseline-data}) and so we use a year's worth of daily observations of (X,Y, Z) values.  Initially, 50 (0.05\%) diseased individuals are introduced to the population of 100,000 individuals (approximately the average population size of a US county in 2015).  The standard deviation of the X values is 0.1\% of the magnitude of the initial number of susceptible individuals, and the standard deviation of the Y values is 10\% of the magnitude of the initial number of infectious individuals, although in this setting the number of infectious individuals increases rapidly.  We use Gaussian error as our baseline error because of its known statistical properties so we can more easily compare methods to one another.  Later on, we explore the autoregressive and PAVA simulated data, which may better simulate data from  a real epidemic.

      \subsection{Other Simulated Data}

We then generate different SIR data sets, changing one of the conditions for simulation each time and using the baseline SIR data as our basis for comparison. The different data sets generated are shown in Table \ref{tab:simulated-data}, for a total of 27 simulated data sets.

% \begin{table}[h]
% 	\centering
% 	\resizebox{\textwidth}{!}{
% 		\begin{tabular}{@{}lll@{}}
% 			\toprule
%                   \textbf{Parameter Values ($\beta, \gamma$)}        & \textbf{Starting Compartment Sizes ($X_0, Y_0$)}            & \textbf{Variance of Added Errors ($\sigma_X, \sigma_Y$ )}       \\ \midrule
%                   (0.06, 0.001)            & (99000, 1000)                           & (50, 2)                     \\
%                   (0.06, 0.04)             & (99900, 100)               & (500, 20)     \\
%                   (0.06,.03) & (99950, 50) & (100, 5)\\
%                   (0.06, 0.06)             & (99990, 10)             & (2500, 100)           \\
%                   (0.06, 0.24) & (99999, 1) & (10000, 500)                                      \\ \bottomrule
% 		\end{tabular}
% 	}
% 	\caption{Table of data sets.}
% 	\label{tab:datasets}
% \end{table}

% % Please add the following required packages to your document preamble:
% % \usepackage{booktabs}
% \begin{table}[]
% \centering

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\centering
\begin{tabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{Model} & \textbf{Error} & \textbf{($\beta, \gamma$)} & \textbf{T} & \textbf{(X(0),Y(0))} & \textbf{N} & \textbf{($\sigma_X, \sigma_Y$)} \\ \midrule
  Baseline        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
  AR       & SIR      &   AR ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
  \textcolor{red}{Norm-M}      & SIR      &   Norm-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
   \textcolor{red}{AR-M}       & SIR      &   AR-M ($\rho=.7)$             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ \midrule
  InfRec1       & SIR      &   Norm             & (60000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)\\
  InfRec2      & SIR      &   Norm             & (1500, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
  InfRec3      & SIR      &   Norm             & (1000, .001)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
InfRec4      & SIR      &   Norm             & (06, .24)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
  Time1 & SIR &  Norm   & (.06, .03)                          &     200       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
  Time2 & SIR &  Norm   & (.06, .03)                          &     100       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
  Time3 & SIR &  Norm   & (.06, .03)                          &     50       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\
  Time4 & SIR &  Norm   & (.06, .03)                          &     20       &  (99950, 50)                    &  $10^5$          &(100, 5)                                \\ \midrule
  Initial1    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99999, 1)                    &  $10^5$          &(100, 5)                                \\
  Initial2    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99990, 10)                    &  $10^5$          &(100, 5)                                \\
  Initial3    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99900, 100)                    &  $10^5$          &(100, 5)                                \\
  Initial4    & SIR      &   Norm             & (.06, .03)                          &     365       &  (99000, 1000)                    &  $10^5$          &(100, 5)                                \\ \midrule
  \textcolor{red}{PopSize1}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (999500, 500)                    &  $10^6$          &(100, 5)                                \\
  \textcolor{red}{PopSize2}       & SIR      &   Norm             & (.06, .03)                          &     365       &  (9995, 5)                    &  $10^4$          &(100, 5)                                \\
  \textcolor{red}{PopSize3}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (990, 10)                    &  $10^3$          &(100, 5)                                \\
  \textcolor{red}{PopSize4}        & SIR      &   Norm             & (.06, .03)                          &     365       &  (99, 1)                    &  $10^2$          &(100, 5)                                \\
 \midrule
  Var1      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(50, 2)                                \\
  Var2      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(500, 20)                                \\
  Var3      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(2500, 1000)                                \\
  Var4      & SIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(10000, 500)                                \\ \midrule
  \textcolor{red}{Poly1}     & Linear      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\
  \textcolor{red}{Poly2}    & Quartic      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
   LSIR     & LSIR      &   Norm             & (.06, .03)                          &     365       &  (99950, 50)                    &  $10^5$          &(100, 5)                            \\ 
  \bottomrule
\end{tabular}
\caption{Table of simulated data sets.  We simulate 27 total data sets with the given parameters.}
\label{tab:simulated-data}
\end{table}

To test how well the SIR models perform under misspecification, we simulate data from three other models. While we will not have a ``known'' $\rr$ value to compare our results to, this will allow us to see if the models give us reasonable (and comparable) results when the data do not come from the SIR model. We simulate data from polynomials of degree 1 and 4 with respect to the time step, using a monotonically increasing function for the $X$ compartment and a monotonically decreasing function for the $Z$ compartment.   The sum of these is subtracted from the constant total population to obtain the $Y$ compartment. We call these models polynomial of degree 1 (Poly1) and polynomial of degree 4 (Poly4), respectively. We also generate data from a linearized version (in terms of the ODEs and not to be confused with LMA) of the three compartment SIR model (LSIR) with the following linearized ODEs:

\begin{eqnarray*}
	\frac{dX}{dt} &=& -\beta Y \\
	\frac{dY}{dt} &=& \beta Y - \gamma Y \\
	\frac{dZ}{dt} &=& \gamma Y.
\end{eqnarray*}

% The data from the linear SIR model were simulated under the following conditions:

% \begin{center}

% 	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2

% 	Number of time steps: 365

% 	Starting X size ($X_0$): 99950

% 	Starting Y size ($Y_0$): 50

% 	Starting Z size: 0

% 	$\sigma_X$: 10

% 	$\sigma_Y$: 1

% 	Total population: 100000.

%       \end{center}

\subsection{Model Fitting Details}\label{sec:model-fit}

Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods. The IPR and SIPR models require knowledge of the length of the average infectious period. For the results below, we assume the best case scenario that the true average infectious period is known for the data set.

\begin{table}[H]
	\centering
	\begin{tabular}{@{}ll@{}}
		\toprule
		\textbf{Model} & \textbf{Default Standard Error Method} \\ \midrule
		Least Squares (LS) & Delta method with Hessian matrix\\
		Reparameterized Least Squares (ReLS) & Hessian matrix \\
		Linear Model Approximation (LMA) & Delta method with linear regression standard error \\
		LMA, All Time Points  (LMAT)& Delta method with linear regression standard error \\
		Incidence-Prevalence Ratio (IPR) & Jackknife resampling \\
		Smoothed IPR (SIPR) & Jackknife resampling \\
		Log-Linear (LL) & Linear regression standard error \\
		Markov Chain (MC) & Delta method with GLM regression standard error \\
		Sequential Bayes (SB) & Standard deviation of posterior distribution\\
		% Exponential Growth (EG) & Delta method with linear regression standard error \\
		% IDEA & Delta method with linear regression standard error \\
		\bottomrule
	\end{tabular}
	\caption{Table of methods for standard errors.}
	\label{tab:se-methods}
\end{table}

Table \ref{tab:baseline-res} shows the ways that standard errors are acquired for the different methods. The IPR, SIPR, and IDEA models require knowledge of the length of the average infectious period. For the results below, we assume the best case scenario that the true average infectious period is known for the data set. The exponential growth model requires knowledge of the probability distribution of the infectious period. In this case, we assume a uniform probability distribution with a mean equivalent to the true average infectious period.

\section{Results of Simulations}\label{sec:results}
There are numerous aspects to consider when using a method to estimate $\rr$.  Among these include values of both the infection and recovery rate $(\beta, \gamma)$, the number of data points available ($T$), the total size of the population ($N$), the initial percent of susceptible individuals $\left (\frac{X(0)}{N}\right)$, the initial percent of infectious individuals $\left (\frac{Y(0)}{N}\right )$, the magnitude of the variance of both the number of susceptibles and the number of infectious ($\sigma_X^2, \sigma_Y^2$), the error modeling assumptions, data generated from a non-SIR model, and finally methods of estimating the variance.  We first examine our baseline data, and then we examine each of these aspects individually.

\subsection{Baseline Simulated Data}\label{sec:res-base}
We implement the \wxxsir methods to estimate $\rr$ using the baseline data set.  The results are displayed in Figure \ref{fig:baseline-res} and Table \ref{tab:baseline-res}.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{images/BaseBase.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the baseline data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}\label{fig:baseline-res}
  \end{figure}

\begin{table}[H]	
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9999711 & 0.0056\\
		\hline
		ReLS & 1.9997340 & 0.0050\\
		\hline
		LMA & 1.6863640 & 0.1886\\
		\hline
		LMAT & 1.4063227 & 0.6309\\
		\hline
		IPR & 4.4203632 & 12.3593\\
		\hline
		SIPR & 1.8870824 & $<$ 1e-04 \\
		\hline
		LL & 1.9998072 & 0.0002\\
		\hline
		MC & 0.9485555 &  $<$ 1e-04 \\
		\hline
		SB & 1.6495023 & 0.0672\\
		\hline
	\end{tabular}
        \caption{Baseline Data $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$, $\rr$ Estimates and Standard Errors}\label{tab:baseline-res}
\end{table}

Overall, in Figure \ref{fig:baseline-res}, we see a wide variation in the estimates of and the standard errors for $\rr$ depending on the method used.  The methods of LS, ReLS, LMA, LMAT, IPR, and LL all result in a CI that covers the true value of $\rr=2$.  Of these, LMAT and IPR result in a CI which contains 1, so we would not be able to say whether or not we would have an outbreak of a disease or not. On the other hand, many methods appear to underestimate the true $\rr$, and many of them have larger standard errors as well.  We would say that IPR's CI is uninformative. Additionally, MC and SIPR models seem to have unreasonably small standard errors.

\subsection{Autoregressive Errors}\label{sec:res-AR}
We next examine the estimates of $\rr$ when fit to the autoregressive simulated data with $\rho=0.7$.  In this data, we may see errors accumulate over time, which may more accurately reflect the course of infection in an epidemic.  For example, we may expect to see more variation around the peak of an epidemic than the beginning.  Besides the AR error, the other parameters are the same as in the baseline model.  The results of the estimates are displayed in Figure \ref{fig:ar-res} and Table \ref{tab:ar-res}.

In this setting, the CIs from LS, ReLS, LMAT, SIPR, and LL cover the true $\rr=2$.  In contrast to the baseline data, the CI for LMA no longer covers the true estimate.  For all of these models, the standard error is larger, which we would expect given that the data have a large positive autocorrelation coefficient.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{images/AR.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the autoregressive (AR) ($\rho=.7)$ simulated data set $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
  \end{figure}\label{fig:ar-res}
\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9997 & 0.0072\\
		\hline
		ReLS & 2.0001 & 0.0063\\
		\hline
		LMA & 1.4745 & 0.1130\\
		\hline
		LMAT & 1.3694 & 0.8682\\
		\hline
		IPR & 1.7127 & 5.0713\\
		\hline
		SIPR & 0.8792 & $<$ 1e-04 \\
		\hline
		LL & 2.0000 & 0.0003\\
		\hline
		MC & 0.9486 & $<$ 1e-04 \\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
        \caption{Autoregressive Errors, $\rr$ Estimates and Standard Errors}\label{tab:ar-res}
\end{table}

\subsection{PAVA Monotonicity}\label{sec:res-PAVA}
We examine the case of the PAVA error assumptions which enforces monotonic assumptions on the suceptible and recovered compartments.  This in turn may more accurately model the course of an epidemic, but as a consequence, we may not be able to guarantee that a 95\% CI covers the true value of $\rr$ 95\% of the time.

In Figure \ref{fig:pava-res} and Table \ref{tab:pava-res}, we display the results of fitting the methods on two data sets, those generated from the Norm-G and Norm-AR error assumptions.  We first see that there is very little difference among the estimates of $\rr$ in both Norm-G and Norm-AR error assumptions.  We see that LMA and IPR have smaller estimates of $\rr$ and SB has a larger estimate of $\rr$.  The standard errors for Norm-AR error assumption are greater than or equal to than the standard error in the Norm-G error assumption with the exceptions of LMAT and SB.

The methods that contain the true value of $\rr$ are LS, ReLS, LMAT, IPR for both error assumptions and LL with Norm-AR error assumption.  The methods LS, ReLS, LMA, and LL would be able to reject the hypothesis that $\rr < 1$.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/mono.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Norm-M and AR-M data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical lines are the true $\rr$ value.}
	\label{fig:pava-res}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r|r|r}
		\hline
		Model & Norm-M Est. & Norm-M SE & AR-M Est. & AR-E SE\\
		\hline
		LS & 2.0002 & 0.0044 & 1.9999 & 0.0074\\
		\hline
		ReLS & 2.0003 & 0.0040 & 1.9997 & 0.0065\\
		\hline
		LMA & 1.6794 & 0.1103 & 1.5350 & 0.1497\\
		\hline
		LMAT & 1.4422 & 0.9242 & 1.4491 & 0.7464\\
		\hline
		IPR & 0.7486 & 1.5047 & 0.7513 & 1.5606\\
		\hline
		SIPR & 0.9082 & $<$ 1e-04 & 0.6976 & $<$ 1e-04\\
		\hline
		LL & 2.0003 & 0.0001 & 1.9996 & 0.0002\\
		\hline
		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523 & 1.7143505 & 0.0685\\
		\hline
	\end{tabular}
\caption{Gaussian and Autoregressive Data with PAVA Applied, $\rr$ Estimates and Standard Errors}\label{tab:pava-res}
\end{table}

In comparison to the baseline data set, the estimates of $\rr$ from the methods fit on the Norm-G error assumption data set have smaller CIs.  The estimates of $\rr$ for the Norm-G error assumption are very similar to those from the baseline assumption.  LMA and LMAT have estimates closer to the true $\rr$ in the Norm-G error assumption data setting compared to the baseline data and SB has an estimate further away.

Overall, there is not much difference between the baseline and monotone data sets, at least in the case with $\sigma_X=100$ and $\sigma_Y=5$.

\subsection{Infection and Recovery Rate ($\beta, \gamma$)}\label{sec:res-beta-gamma}

We next examine the simulated data sets InfRec1-4, three of which have different levels of $\gamma$ for a given level of $\beta = 0.06$ and the fourth is simulated with $\rr=0.25 < 1$, a case where an outbreak would not occur.  The idea behind this is that since $\rr=\frac{\beta}{\gamma}$ and ratio estimators typically have large variance, we want to examine how a small value of $\gamma$, that is a small recovery rate, would effect the resulting estimates of $\rr$.  Additionally, we would like to know when a disease will die out on its own accord, that we can be certain the estimate of $\rr$ is less than 1.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{images/parchange_n.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
  \label{fig:infrec1-res}
\end{figure}
\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		SIRLS & 59.9765 & 0.0435\\
		\hline
		ReLS & 59.9952 & 0.4344\\
		\hline
		LMA &  -117.9695 & 101.1886 \\
		\hline
		LMAT & -2.8536 & 193.6526 \\
		\hline
		IPR & 54.1465 & 200.2595\\
		\hline
		SIPR & 3.9645 & $<$ 1e-04\\
		\hline
		LL & 46.3210 & 0.7797\\
		\hline
		MC & 0.9486 & $<$ 1e-04 \\
		\hline
		SB & 0.9767 & 0.0517\\
		\hline
	\end{tabular}
        \caption{\label{tab:infrec1-res}$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
\end{table}

Looking in Figure \ref{fig:infrec1-res} and Table \ref{tab:infrec1-res} at the cases where $\beta = 0.06$ with $\gamma = 0.001, 0.04, 0.01, 0.24$, it appears that most of the models do not handle a small $\gamma$ well.  We examine an exaggerated extreme case of when $\rr=60$ (the largest reported $\rr$ value we found is $\hat{\rr}=14$ for measles as reported in \cite{anderson1992}). Only LS and ReLS result in CIs that cover the true $\rr$. IPR and LL give fairly close point estimates to the true $\rr=60$ as well. However, the other models do very poorly; they tend to greatly underestimate $\rr$. The linear model approximation does particularly poorly, likely because the linear model does not fit the data well, and is not even constrained to positive values of $X$, $Y$, and $Z$, which is why we can result in a negative estimate of $\rr$. Several of the models also return extremely large standard errors, meaning they are not very reliable in these cases.  Only LS, ReLS, and LL would reject the hypothesis that $\rr <1$, when this is clearly not the case at all.

When $\rr=1.5$ we see that the estimates from the methods give results much closer to the true value, and the resulting CIs are also smaller.  The methods LS, ReLS, SIPR, LL, and SB would all reject the hypothesis that $\rr > 1$.  Although if we are testing the null hypothesis for  all methods simultaneously, we should use a multiple-testing correction such as Bonferroni.  Under the Bonferroni correction in this case, the only method which may then fail to reject would be SB as the other SEs are very small).

When $\rr=1$, we are unsure whether there will be an epidemic or not.  In this case, all the CIs of the methods cover the true value.  We also see that the SE of LS is much greater than ReLS.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 0.2363 & 4.1043 \\
		\hline
		ReLS & 4.1795 & 2.7456\\
		\hline
		LMA &  0.7075 & 0.2029 \\
		\hline
		LMAT & 0.6747 & 6.2471 \\
		\hline
		IPR & 181.0158 & 742.1088 \\
		\hline
		SIPR & -0.2484 & 0.1160 \\
		\hline
		LL & 0.9912 & 0.0031\\
		\hline
		MC & 0.9486 & 0.0001\\
		\hline
		SB & 1.2355 & 0.0582\\
		\hline
	\end{tabular}
        \caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}\label{tab:infrec4-res}
\end{table}

Table \ref{tab:infrec4-res} corresponds to the case where $\gamma = 0.24$ and $\beta=0.06$. It appears that a lot of estimation methods do not handle small $\rr$ well either. None of them are extremely close to truth, except for ReLS, as most of the tend to overestimate the truth. The IPR is particularly bad in this case. Also, in general, the standard errors are several orders of magnitude larger than in the baseline data.

\subsubsection{Gaussian Monotonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/parchange_nm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}
\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 59.9765 & 0.0387\\
		\hline
		ReLS & 59.9952 & 0.3875\\
		\hline
		LMA & -55.3749 & 10.9872\\
		\hline
		LMAT & 7.7686 & 103.2206\\
		\hline
		IPR & 24.4087 & 66.1668\\
		\hline
		SIPR & 3.8285 & $<$ 1e-04\\
		\hline
		LL & 59.8300 & 0.1079\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 0.2328 & 0.6701\\
		\hline
		ReLS & 3.5710 & 0.3171\\
		\hline
		LMA & 0.5217 & 0.1716\\
		\hline
		LMAT & 0.3872 & 4.8818\\
		\hline
		IPR & 6.6728 & 89.9362\\
		\hline
		SIPR & -0.0317 & 0.0336\\
		\hline
		LL & 0.8623 & 0.0088\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.5587 & 0.0653\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/parchange_ar.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}
\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 60.0151 & 0.0698\\
		\hline
		ReLS & 60.0255 & 0.6912\\
		\hline
		LMA & -19.2058 & 5.9751\\
		\hline
		LMAT & 76.3404 & 472.4155\\
		\hline
		IPR & 58.7237 & 219.9891\\
		\hline
		SIPR & 3.9056 & $<$ 1e-04\\
		\hline
		LL & 41.8977 & 0.9768\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & -0.0695 & 2.8291\\
		\hline
		ReLS & 5.1930 & 3.5250\\
		\hline
		LMA & 0.8750 & 0.0591\\
		\hline
		LMAT & 1.7865 & 15.9420\\
		\hline
		IPR & 52.5000 & 740.0182\\
		\hline
		SIPR & -2.8643 & $<$ 1e-04\\
		\hline
		LL & 1.0014 & 0.0025\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0160 & 0.0528\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Monotonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/parchange_arm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the infrec1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}
\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 59.9589 & 0.0631\\
		\hline
		ReLS & 59.9961 & 0.6215\\
		\hline
		LMA & -183.9423 & 277.8292\\
		\hline
		LMAT & -69.9068 & 1603.7792\\
		\hline
		IPR & 28.5231 & 120.9288\\
		\hline
		SIPR & 3.7113 & $<$ 1e-04\\
		\hline
		LL & 70.7076 & 0.9517\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.001$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 0.3021 & 0.5657\\
		\hline
		ReLS & 9.8415 & 1.7074\\
		\hline
		LMA & -49.7813 & 5303.0302\\
		\hline
		LMAT & 0.2221 & 5.4112\\
		\hline
		IPR & 0.8145 & 11.8709\\
		\hline
		SIPR & 0.7365 & 0.0675\\
		\hline
		LL & 0.9199 & 0.0053\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$\gamma = 0.24$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsection{Number of Time Points ($T$)}\label{sec:res-time}
We examine how the number of time steps the estimates of $\rr$ in the different methods.  Ideally, the more data we have, the more likely we are to have lower-variance estimates for $\rr$.  Since we would like to  infer how a disease will spread through a population, the sooner we can be sure whether $\rr> 1$, the sooner we can implement interventions and control measures for a disease.  Another reason to analyze the number of time steps is that some of the methods assume linear or exponential growth of an epidemic in a certain time period.  If the growth is no longer exponential, we will obtain a poor estimate of $\rr$.  In this case, we can be using too many time points above a certain time.

In Figure \ref{fig:time-res} and Tables \ref{tab:time-res1} and \ref{tab:time-res2} we examine the methods fit to different amounts of time periods where we observe the progress of a disease, including $T=200$, 100, 50, and 20.

Again, we see that IPR does very poorly in estimating $\rr$, regardless of the amount of data observed.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{images/time_n.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}\label{fig:time-res}
  \end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9993 & 0.0286\\
		\hline
		ReLS & 1.9994 & 0.0313\\
		\hline
		LMA & 0.9496 & 0.0825\\
		\hline
		LMAT & 3.3443 & 21.4895\\
		\hline
		IPR & 7.3458 & 16.1289\\
		\hline
		SIPR & -2.1199 & $<$ 1e-04\\
		\hline
		LL & 2.0014 & 0.0018\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 2.4289 & 0.1102\\
		\hline
	\end{tabular}
        \caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}\label{tab:time-res1}
\end{table}

Table \ref{tab:time-res1} shows the methods applied to the first 100 time steps, out of 365, of the baseline data set . The estimates and standard errors are worse than baseline, which we would expect given less data. The results are still relatively close to the full baseline data. There also appears to be less tendency to underestimate $\rr$ compared to a longer time scale.  In general, we see that that the length of CIs decreases as we use more and more time points.  Only LL would reject the hypothesis that $\rr < 1$ for 50 or fewer time points, but its CI does not cover the true $\rr$ value.  With 50 and greater time points, we see that LS and ReLS provide close point estimates to the true value and by 100 time points, would reject the hypothesis that $\rr < 1$.  The methods and MC and SB consistently have small CIs, but never cover the true value.

\begin{table}[H]
  \centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 38.6896 & 106.2705\\
		\hline
		ReLS & 22.9646 & 6.9817\\
		\hline
		LMA & 1.4418 & 7.3164\\
		\hline
		LMAT & 1.0079 & 0.1425\\
		\hline
		IPR & 25.7313 & 35.2926\\
		\hline
		SIPR & 4.1772 & 7.3203\\
		\hline
		LL & 1.0430 & 0.0276\\
		\hline
		MC & 0.9486 & 5.0e-07\\
		\hline
		SB & 1.0000 & 0.2236\\
		\hline
	\end{tabular}
        \caption{$T$ = 20, $\rr$ Estimates and Standard Errors}\label{tab:time-res2}
\end{table}

Table \ref{tab:time-res2} shows the methods applied to the first 20 time steps of a baseline data set. As might be expected, the $\rr$ estimates are wildly inaccurate, and none are very close to the true value. The methods of LS and ReLS are particularly worse in this situation compared to the other data sets. The standard errors are much larger than the baseline, which is to be expected as well due to the small number of data points. This shows that it is extremely difficult to create accurate online estimates of $\rr$ or to estimate $\rr$ without adequate data.

\subsubsection{Gaussian Monotonic Errors}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/time_nm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9934 & 0.1464\\
		\hline
		ReLS & 1.9935 & 0.1975\\
		\hline
		LMA & 1.2024 & 0.5439\\
		\hline
		LMAT & 2.7981 & 3.8629\\
		\hline
		IPR & 2.2784 & 5.3298\\
		\hline
		SIPR & 2.2020 & 0.5643\\
		\hline
		LL & 1.9928 & 0.0105\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.1000\\
		\hline
	\end{tabular}
	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 5.0366 & 1.3672\\
		\hline
		ReLS & 5.0449 & 2.1398\\
		\hline
		LMA & 1.1218 & 1.7267\\
		\hline
		LMAT & 2.7046 & 4.8031\\
		\hline
		IPR & 1.4155 & 4.5655\\
		\hline
		SIPR & 0.9703 & 0.8674\\
		\hline
		LL & 1.7472 & 0.3635\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.2236\\
		\hline
	\end{tabular}
	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/time_ar.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9098 & 0.3811\\
		\hline
		ReLS & 1.9104 & 0.4773\\
		\hline
		LMA & 0.9141 & 0.2563\\
		\hline
		LMAT & 0.7524 & 2.5623\\
		\hline
		IPR & 9.2219 & 18.1185\\
		\hline
		SIPR & 2.0919 & 1.0699\\
		\hline
		LL & 1.7390 & 0.0358\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 3.3635 & 0.1834\\
		\hline
	\end{tabular}
	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 0.0099 & 2.1688\\
		\hline
		ReLS & -0.1539 & 2.7692\\
		\hline
		LMA & -5.2984 & 65.6503\\
		\hline
		LMAT & 0.5717 & 1.3982\\
		\hline
		IPR & 20.4892 & 33.7260\\
		\hline
		SIPR & 0.3570 & 7.6611\\
		\hline
		LL & 0.9171 & 0.0338\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 11.3793 & 0.7543\\
		\hline
	\end{tabular}
	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Monotonic Errors}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/time_arm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Time1-4 data sets with Gaussian errors $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9674 & 0.2324\\
		\hline
		ReLS & 1.9679 & 0.3065\\
		\hline
		LMA & 1.1584 & 0.3660\\
		\hline
		LMAT & 3.4346 & 14.6041\\
		\hline
		IPR & 2.2454 & 5.7090\\
		\hline
		SIPR & 1.9631 & 0.3489\\
		\hline
		LL & 2.0524 & 0.0209\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.1000\\
		\hline
	\end{tabular}
	\caption{ $T$ = 100, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]
	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.6816 & 0.4747\\
		\hline
		ReLS & 1.6813 & 0.6544\\
		\hline
		LMA & 0.8899 & 0.7605\\
		\hline
		LMAT & 1.4496 & 1.5423\\
		\hline
		IPR & 3.8856 & 8.2899\\
		\hline
		SIPR & 3.9849 & 2.1351\\
		\hline
		LL & 1.4328 & 0.0397\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.2236\\
		\hline
	\end{tabular}
	\caption{$T$ = 20, $\rr$ Estimates and Standard Errors}
\end{table}

\subsection{Initial Percentage of Infectious $\left (\frac{Y(0)}{N}\right)$}\label{sec:res-inf}
We next examine the initial conditions of the number of susceptible, infectious, and recovered.  Deterministically, different initial conditions in the number of infectious can cause very different epidemic courses (see Figures in appendix).

In Figure \ref{fig:inits-res} and Tables \ref{tab:inits-res1} and \ref{tab:inits-res2}, we display the results of different number of initially infectious individuals of 1, 10, 100, and 1000 with a constant population of 100,000.  We see that for a smaller percentage of initially infectious individuals, we have larger CIs in our estimates of $\rr$.  In fact, LL is the only method that has a consistently small CI around the true $\rr=2$, regardless of the starting number of infectious.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{images/start_n.jpeg}
    \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}\label{fig:inits-res}
  \end{figure}

  Table \ref{tab:inits-res1} shows the methods applied to data where $Y(0) = 1000$, essentially starting at the middle of an epidemic. This data has very large underestimation of $\rr$. The LMA method once again gives an estimate very far from the true $\rr$, potentially indicating that a linear approximation does not apply in this case. The sample standard errors in this case are actually relatively close to the baseline, and even smaller for some of the methods.

Table \ref{tab:inits-res2} shows the methods applied to data where $Y(0) = 1$, starting at the very beginning of an epidemic, and the true technical definition of $\rr$ where we introduce exactly one infectious individual into a population. The methods give many larger estimates, tending to overestimate the true $\rr$. In particular, IPR and LMA methods give very large results. We also see a larger standard error than the baseline.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9996 & 0.0050\\
		\hline
		ReLS & 1.9998 & 0.0034\\
		\hline
		LMA & -87.5946 & 66413.1280\\
		\hline
		LMAT & 0.7306 & 4.6810\\
		\hline
		IPR & 1.2851 & 1.4410\\
		\hline
		SIPR & 1.2522 & 57.9230\\
		\hline
		LL & 1.9996 & 0.0003 \\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.9729 & 0.0516\\
		\hline
	\end{tabular}
        \caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res1}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 2.4683 & 1.5371\\
		\hline
		ReLS & 2.4692 & 2.7294 \\
		\hline
		LMA & 10.9282 & 19.2974 \\
		\hline
		LMAT & 1.6246 & 5.0360\\
		\hline
		IPR & 774.2747 & 7491.0051 \\
		\hline
		SIPR & 2.4391 & $<$ 1e-04 \\
		\hline
		LL & 1.9997 & 0.0008\\
		\hline
		MC & 0.9486 & 1.56e-05\\
		\hline
		SB & 1.4715 & 0.0635\\
		\hline
	\end{tabular}
        \caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}\label{tab:inits-res2}
\end{table}

\subsubsection{Gaussian Montonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/start_nm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0003 & 0.0046\\
		\hline
		ReLS & 2.0003 & 0.0032\\
		\hline
		LMA & -0.9982 & 1.4189\\
		\hline
		LMAT & 0.9703 & 0.6574\\
		\hline
		IPR & 0.9866 & 1.0570\\
		\hline
		SIPR & 2.0166 & 57.8154\\
		\hline
		LL & 2.0001 & 0.0002\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0868 & 0.0546\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.4668 & 1.5349\\
		\hline
		ReLS & 2.4670 & 2.7219\\
		\hline
		LMA & -99.3726 & 3136.3141\\
		\hline
		LMAT & 0.4205 & 12.9331\\
		\hline
		IPR & 3.6061 & 26.5409\\
		\hline
		SIPR & 5.4116 & $<$ 1e-04\\
		\hline
		LL & 2.0005 & 0.0004\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.7685 & 0.0696\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/start_ar.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0005 & 0.0078\\
		\hline
		ReLS & 2.0001 & 0.0054\\
		\hline
		LMA & 4.1646 & 3.8688\\
		\hline
		LMAT & 3.1508 & 45.5885\\
		\hline
		IPR & 1.1936 & 1.1958\\
		\hline
		SIPR & 2.2854 & 57.4383\\
		\hline
		LL & 1.9998 & 0.0004\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.3154 & 0.0294\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.4706 & 1.5372\\
		\hline
		ReLS & 2.4707 & 2.7245\\
		\hline
		LMA & -0.8259 & 1.7277\\
		\hline
		LMAT & 1.3117 & 26.0563\\
		\hline
		IPR & 5.0000 & 1199.1557\\
		\hline
		SIPR & 1.2047 & $<$ 1e-04\\
		\hline
		LL & 2.0027 & 0.0011\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.7698 & 0.0696\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Montonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/start_arm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Inits1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, \sigma_X=100, \sigma_Y=5, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0001 & 0.0069\\
		\hline
		ReLS & 2.0000 & 0.0048\\
		\hline
		LMA & -10.8054 & 286.9861\\
		\hline
		LMAT & 0.9333 & 0.8824\\
		\hline
		IPR & 0.9960 & 0.9987\\
		\hline
		SIPR & 1.9222 & 57.0458\\
		\hline
		LL & 2.0006 & 0.0003\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.3790 & 0.0322\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.4667 & 1.5349\\
		\hline
		ReLS & 2.4665 & 2.7349\\
		\hline
		LMA & 16.7344 & 18.2294\\
		\hline
		LMAT & 3.1984 & 12.9100\\
		\hline
		IPR & 2.5082 & 12.8639\\
		\hline
		SIPR & 1.0356 & $<$ 1e-04\\
		\hline
		LL & 1.9999 & 0.0006\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.7793 & 0.0698\\
		\hline
	\end{tabular}
	\caption{$Y(0) = 1$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsection{Population Size $N$}\label{sec:res-n}
We next examine the effects on the estimates of $\rr$ when changing the total population size, $N$.  The percent of initially infectious individuals is 1\% when $N=10,000$ and $N=100,000$ and .5\% when $N=1,000$ and $N=100$. Intuitively, we would expect to see the variance decrease and to have more accurate estimates with a larger $N$ compared to a smaller one, due to the central limit theorem.  This is exactly what we see for some of the methods, but not all of them.

We display the results of estimating $\rr$ from the methods fit on the baseline data set with varying $N$ in Figure \ref{fig:inits-res} and Tables \ref{tab:n1-res2} and \ref{tab:n2-res2}.  When $N=100$, LS, ReLS, LMA, LMAT, IPR, and SIPR's 95\% CIs all cover the true $\rr$.  However, we would say that all of these intervals are uninformative.  We see that CIs of LS are wider than those of ReLS, although the difference becomes less substantial as $N$ grows larger.

When $N=100$, we would be able to say the estimate for $\rr$ from LL is greater than 1 with 95\% confidence, but the estimate is much smaller than the true value.  MC and SB would actually allow us to say $\rr < 1$ with 95\% confidence.  It is not until $N=10,000$ that we are able to say with certainty that LS and ReLS, in addition to LL, give informative estimates of $\rr$.

Thus, it seems that if one is estimating for $\rr$ for a small population such as a small village or crusie ship, LL may be the most appropriate method, but when the population is larger (~10,000), we can be confident in using LS or ReLS.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/popsize_n.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}\label{fig:inits-res2}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 2.0008 & 0.0590\\
		\hline
		ReLS & 2.0005 & 0.0528\\
		\hline
		LMA & 1.9976 & 2606.0798\\
		\hline
		LMAT & 1.2944 & 1.4305\\
		\hline
		IPR & 40.8072 & 133.2798\\
		\hline
		SIPR & 1.0629 &  $<$ 1e-04\\
		\hline
		LL & 1.9988 & 0.0023\\
		\hline
		MC & 0.9486 & 4.0e-06\\
		\hline
		SB & 0.4418 & 0.0348\\
		\hline
	\end{tabular}
\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}\label{tab:n1-res2}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9863 & 4.8620\\
		\hline
		ReLS & 1.9871 & 3.4569\\
		\hline
		LMA & 0.8023 & 18.9417\\
		\hline
		LMAT & 1.2209 & 4.7699\\
		\hline
		IPR & 3620.8677 & 57608.1700\\
		\hline
		SIPR & 3.3678 & 22.4187\\
		\hline
		LL & 1.1413 & 0.0422\\
		\hline
		MC & 0.8456 & 0.0125\\
		\hline
		SB & 0.3887 & 0.0326\\
		\hline
	\end{tabular}
\caption{$N = 100$, $\rr$ Estimates and Standard Errors}\label{tab:n2-res2}
\end{table}

\subsubsection{Gaussian Monotonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/popsize_nm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0175 & 0.1430\\
		\hline
		ReLS & 2.0175 & 0.1021\\
		\hline
		LMA & 0.8256 & 1.4444\\
		\hline
		LMAT & 1.0457 & 0.6365\\
		\hline
		IPR & 1.7343 & 12.7260\\
		\hline
		SIPR & 1.6932 & 12.0122\\
		\hline
		LL & 2.0102 & 0.0056\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9998 & 0.0006\\
		\hline
		ReLS & 1.9999 & 0.0005\\
		\hline
		LMA & 1.7673 & 0.0976\\
		\hline
		LMAT & 1.9732 & 10.8872\\
		\hline
		IPR & 1.4344 & 1.1311\\
		\hline
		SIPR & 1.7907 & $<$ 1e-04\\
		\hline
		LL & 2.0000 & $<$ 1e-04\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.2709 & 0.0590\\
		\hline
	\end{tabular}
	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/popsize_ar.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.1221 & 0.7363\\
		\hline
		ReLS & 2.1219 & 0.5660\\
		\hline
		LMA & 1.2045 & 0.3878\\
		\hline
		LMAT & 1.0881 & 3.9239\\
		\hline
		IPR & 164.6111 & 1887.1253\\
		\hline
		SIPR & 0.5335 & 13.6519\\
		\hline
		LL & 2.4701 & 0.0546\\
		\hline
		MC & 0.9500 & 0.0003\\
		\hline
		SB & 1.5457 & 0.0651\\
		\hline
	\end{tabular}
	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9998 & 0.0009\\
		\hline
		ReLS & 2.0001 & 0.0007\\
		\hline
		LMA & 1.6873 & 0.0810\\
		\hline
		LMAT & 1.3885 & 0.6855\\
		\hline
		IPR & 1.5672 & 1.5944\\
		\hline
		SIPR & 1.7903 & $<$ 1e-04\\
		\hline
		LL & 1.9999 & $<$ 1e-04\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.1877 & 0.0570\\
		\hline
	\end{tabular}
	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Monotonic Errors}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{images/popsize_arm.jpeg}
	\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the PopSize1-4 data sets $(\beta=.06, \gamma=.03, T=365, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0975 & 0.2921\\
		\hline
		ReLS & 2.0976 & 0.2211\\
		\hline
		LMA & 1.1720 & 0.0511\\
		\hline
		LMAT & -13.6906 & 281.9114\\
		\hline
		IPR & 1.7077 & 11.7775\\
		\hline
		SIPR & 2.8981 & 284.4933\\
		\hline
		LL & 2.2806 & 0.0230\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$N = 10000$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0000 & 0.0007\\
		\hline
		ReLS & 1.9997 & 0.0007\\
		\hline
		LMA & 1.8332 & 0.1199\\
		\hline
		LMAT & 1.4164 & 0.6348\\
		\hline
		IPR & 1.4490 & 1.1803\\
		\hline
		SIPR & 1.7505 & $<$ 1e-04\\
		\hline
		LL & 1.9998 & $<$ 1e-04\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.8078 & 0.0704\\
		\hline
	\end{tabular}
	\caption{$N = 100$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsection{Magnitude of Variance $(\sigma^2_X, \sigma^2_Y)$}\label{sec:res-var}
We examine the estimates of $\rr$ when changing the values of the variance of the number of susceptibles and number of infectious.  We would expect to see wider CIs when we have larger variances in the $X$ and $Y$ compartments, but we would like to determine how sensitive estimates of $\rr$ are to these variance values.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
\begin{center}
  \includegraphics[scale=0.5]{images/var_n.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
  \label{fig:mag-var-res}
\end{center}
\end{figure}

We display the results in Figure \ref{fig:mag-var-res} and Tables \ref{tab:mag-var-res1}-\ref{tab:mag-var-res2}.  Generally, we see wider CIs and more biased estimates of $\rr$ when the variance levels are large.  In particular, LMA and LMAT seem to be particularly effected as the CIs grow substantially as the variance grows.  Again we see that LS, ReLS, and LL provide the most accurate estimates with relatively small CIs.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9998 & 0.0027\\
		\hline
		ReLS & 1.9998 & 0.0024\\
		\hline
		LMA & 2.0398 & 0.2774\\
		\hline
		LMAT & 1.4149 & 0.5780\\
		\hline
		IPR & 2.7608 & 8.1319\\
		\hline
		SIPR & 1.8368 & $<$ 1e-04\\
		\hline
		LL & 1.9999 & 0.0001\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.6750 & 0.0677\\
		\hline
	\end{tabular}
        	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res1}
\end{table}

In Table \ref{tab:mag-var-res1}, the magnitude of the variance of the noise is smaller than in the baseline. The estimates are closer to the true $\rr$ and the standard errors are smaller across the board. This indicates that the methods give more accurate and precise estimates when the data are closer to following underlying true SIR model.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.9981 & 0.1354\\
		\hline
		ReLS & 1.9981 & 0.1208\\
		\hline
		LMA & 0.4652 & 7.5508\\
		\hline
		LMAT & 1.2878 & 4.3948\\
		\hline
		IPR & 81.4902 & 348.7351\\
		\hline
		SIPR & 1.0606 & $<$ 1e-04\\
		\hline
		LL & 2.0019 & 0.0054\\
		\hline
		MC & 0.9486 & 1.0e-06\\
		\hline
		SB & 1.8567 & 0.0713\\
		\hline
	\end{tabular}
        	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}\label{tab:mag-var-res2}
\end{table}

In Table \ref{tab:mag-var-res2}, the magnitude of the variance of the noise is larger than in the baseline. The estimates are worse and the standard errors are larger compared to the baseline data set, which we might expect. These estimates are not as terrible as we might imagine given the magnitude of the change - it seems that these methods are somewhat more robust to large noise compared to other changes. One exception in the IPR model, which may do worse due to larger fluctuations in the incidence prevalence ratio.

\subsubsection{Gaussian Monotonic Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/var_nm.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9998 & 0.0026\\
		\hline
		ReLS & 1.9999 & 0.0024\\
		\hline
		LMA & 1.6268 & 0.0865\\
		\hline
		LMAT & 1.5099 & 2.1244\\
		\hline
		IPR & 1.4509 & 2.0522\\
		\hline
		SIPR & 1.7778 & $<$ 1e-04\\
		\hline
		LL & 2.0000 & 0.0001\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.2333 & 0.0581\\
		\hline
	\end{tabular}
	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9970 & 0.0541\\
		\hline
		ReLS & 1.9971 & 0.0482\\
		\hline
		LMA & 4.2062 & 66.2206\\
		\hline
		LMAT & 1.3653 & 3.2675\\
		\hline
		IPR & 2.2672 & 17.0580\\
		\hline
		SIPR & 1.2346 & $<$ 1e-04\\
		\hline
		LL & 1.9957 & 0.0015\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/var_ar.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9996 & 0.0038\\
		\hline
		ReLS & 1.9999 & 0.0034\\
		\hline
		LMA & 1.4644 & 0.0750\\
		\hline
		LMAT & 1.3999 & 0.7194\\
		\hline
		IPR & 2.1980 & 5.4569\\
		\hline
		SIPR & 1.7604 & $<$ 1e-04\\
		\hline
		LL & 1.9991 & 0.0001\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.6378 & 0.0670\\
		\hline
	\end{tabular}
	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9997 & 0.1894\\
		\hline
		ReLS & 2.0002 & 0.1680\\
		\hline
		LMA & 1.0392 & 0.0166\\
		\hline
		LMAT & 1.1985 & 1.6107\\
		\hline
		IPR & 195.3708 & 1451.0401\\
		\hline
		SIPR & 1.7028 & $<$ 1e-04\\
		\hline
		LL & 2.0386 & 0.0082\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.9073 & 0.0723\\
		\hline
	\end{tabular}
	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Monotonic Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/var_arm.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Var1-4 data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.9990 & 0.0036\\
		\hline
		ReLS & 1.9995 & 0.0032\\
		\hline
		LMA & 2.8515 & 0.5541\\
		\hline
		LMAT & 1.4263 & 0.6782\\
		\hline
		IPR & 1.4465 & 2.1821\\
		\hline
		SIPR & 1.8423 & $<$ 1e-04\\
		\hline
		LL & 1.9997 & 0.0001\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.5555 & 0.0653\\
		\hline
	\end{tabular}
	\caption{ $(\sigma^2_X, \sigma^2_Y) = (50, 2)$, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 2.0021 & 0.0928\\
		\hline
		ReLS & 2.0022 & 0.0827\\
		\hline
		LMA & -0.1164 & 1.9126\\
		\hline
		LMAT & 0.9424 & 4.3914\\
		\hline
		IPR & 2.9241 & 18.4344\\
		\hline
		SIPR & 2.5962 & $<$ 1e-04\\
		\hline
		LL & 2.0157 & 0.0031\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
	\caption{$(\sigma^2_X, \sigma^2_Y) = (2500, 100)$, $\rr$ Estimates and Standard Errors}
\end{table}

\subsection{Other Models}\label{sec:res-oth}
Until this point, we have only examined results while assuming that the epidemic truly follows a SIR model.  However, this may not be the case.  We examine the estimates of $\rr$ under different underlying models including a polynomial of degree 1, a polynomial of degree 4, and a linearized SIR model.  We present the results of the estimates of $\rr$ on these models in Figure \ref{fig:other-res} and Tables \ref{tab:other-res1} - \ref{tab:other-res3}.

\subsubsection{Gaussian Errors}

\begin{figure}[H]
  \begin{center}
    \includegraphics[scale=0.5]{images/other_n.jpeg}
  \caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
  \label{fig:other-res}	
  \end{center}
\end{figure}

For both the Poly1 and LSIR data sets, all the methods besides IPR and SIPR underestimate $\rr$.  Moreover, only the CI of IPR covers that of the true value of $\rr=2$ and the CIs are uninformative.  For the quartic model, only ReLS has a CI that both covers the true $\rr$ while rejecting the hypothesis that $\rr < 1$.  LS covers the true value with its CI but also includes $\rr=1$.  LL rejects the hypothesis that $\rr <$ but its CI does not contain $\rr=2$.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.0514 & 0.3783\\
		\hline
		ReLS & 1.0514 & 0.0259\\
		\hline
		LMA & 1.2656 & 0.1278\\
		\hline
		LMAT & 1.1597 & 0.1020\\
		\hline
		IPR & 4.5000 & 15.2186\\
		\hline
		SIPR & 98.0611 & $<$ 1e-04\\
		\hline
		LL & 1.1881 & 0.0006\\
		\hline
		MC & 0.9486 & 1e-07 \\
		\hline
		SB & 1.4150 & 0.0622\\
		\hline
	\end{tabular}
	\caption{Linear Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res1}
\end{table}

\begin{table}[H]

\centering
\begin{tabular}[t]{l|r|r}
	\hline
	Model & Estimate & Std. Dev\\
	\hline
	LS & 1.7693 & 0.4227\\
	\hline
	ReLS & 1.7693 & 0.2797\\
	\hline
	LMA & 0.3618 & 0.1231\\
	\hline
	LMAT & 0.8213 & 6.0466\\
	\hline
	IPR & 10.4349 & 33.8953 \\
	\hline
	SIPR & 4.7477 &  $<$ 1e-04\\
	\hline
	LL & 1.9552 & 0.0165\\
	\hline
	MC & 0.9486 & 6e-07\\
	\hline
	SB & 0.2119 & 0.02409\\
	\hline
\end{tabular}
\caption{4th Order Model, $\rr$ Estimates and Standard Errors}\label{tab:other-res2} 
\end{table}

Table \ref{tab:other-res2} shows the methods applied to data from a 4th order linear regression model with respect to time. Obviously, there is no true $\rr$ to compare to. What interests us in this case is whether the methods give similar results, since all the estimation methods are derived from the same model. However, the results vary much more compared to baseline data estimates, where underlying data came from SIR. We also get massive standard errors, even though actual noise variance magnitudes are the same as baseline.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Err\\
		\hline
		LS & 1.2743 & 0.0980\\
		\hline
		ReLS & 1.2743 & 0.0340\\
		\hline
		LMA & 1.2191 & 0.1093\\
		\hline
		LMAT & 1.2009 & 0.0031\\
		\hline
		IPR & 2.5280 & 2.5904\\
		\hline
		SIPR & 2.1568 & 0.3585\\
		\hline
		LL & 1.2586 & 0.0010\\
		\hline
		MC & 0.9486 & $<$ 1e-04 \\
		\hline
		SB & 1.2009 & 0.0574\\
		\hline
	\end{tabular}
	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}\label{tab:other-res3}
\end{table}

\subsubsection{Gaussian Monotonic Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/other_nm.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.0514 & 0.3785\\
		\hline
		ReLS & 1.0514 & 0.0259\\
		\hline
		LMA & 1.2300 & 0.0572\\
		\hline
		LMAT & 1.1471 & 0.1751\\
		\hline
		IPR & 2.2972 & 7.1420\\
		\hline
		SIPR & 27.4297 & $<$ 1e-04\\
		\hline
		LL & 1.1880 & 0.0006\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.3647 & 0.0611\\
		\hline
	\end{tabular}
	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.7691 & 0.4216\\
		\hline
		ReLS & 1.7691 & 0.2791\\
		\hline
		LMA & -0.0201 & 0.1128\\
		\hline
		LMAT & 0.9972 & 2.5752\\
		\hline
		IPR & 5.9240 & 17.8114\\
		\hline
		SIPR & 15.0836 & $<$ 1e-04\\
		\hline
		LL & 1.9553 & 0.0164\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.2581 & 0.0266\\
		\hline
	\end{tabular}
	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.2739 & 0.0977\\
		\hline
		ReLS & 1.2741 & 0.0337\\
		\hline
		LMA & 1.2068 & 0.0481\\
		\hline
		LMAT & 1.2002 & 0.0041\\
		\hline
		IPR & 1.2086 & 1.1254\\
		\hline
		SIPR & 1.2944 & 0.2158\\
		\hline
		LL & 1.2587 & 0.0010\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.3209 & 0.0602\\
		\hline
	\end{tabular}
	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/other_ar.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.0519 & 0.3775\\
		\hline
		ReLS & 1.0518 & 0.0260\\
		\hline
		LMA & 1.2766 & 0.1573\\
		\hline
		LMAT & 1.1558 & 0.0687\\
		\hline
		IPR & 4.3348 & 41.3251\\
		\hline
		SIPR & 1.4988 & $<$ 1e-04\\
		\hline
		LL & 1.1876 & 0.0006\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.4979 & 0.0641\\
		\hline
	\end{tabular}
	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.7690 & 0.4228\\
		\hline
		ReLS & 1.7691 & 0.2799\\
		\hline
		LMA & 0.3565 & 0.1259\\
		\hline
		LMAT & 3.2425 & 26.2628\\
		\hline
		IPR & 6.2685 & 20.5362\\
		\hline
		SIPR & 2.7662 & $<$ 1e-04\\
		\hline
		LL & 1.9551 & 0.0165\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.2132 & 0.0242\\
		\hline
	\end{tabular}
	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.2737 & 0.1000\\
		\hline
		ReLS & 1.2742 & 0.0346\\
		\hline
		LMA & 1.2146 & 0.0627\\
		\hline
		LMAT & 1.2032 & 0.0148\\
		\hline
		IPR & 1.3799 & 1.3114\\
		\hline
		SIPR & 1.2967 & 0.2216\\
		\hline
		LL & 1.2589 & 0.0010\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.2039 & 0.0574\\
		\hline
	\end{tabular}
	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
\end{table}

\subsubsection{Autoregressive Monotonic Errors}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/other_arm.jpeg}
		\caption{Forest plots of estimates of $\rr$ for the \xxsir methods for the Poly1, Poly4, and LSIR data sets $(\beta=.06, \gamma=.03, T=365, X(0)=99950, Y(0)=50, Z(0)=0, N=10^5, \sigma_X=100, \sigma_Y=5)$.  The points are the estimates of $\rr$ and the lines denote $\pm 2\cdot $Std. Err.  The vertical line is the true $\rr$ value.}
	\end{center}
\end{figure}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.0513 & 0.3783\\
		\hline
		ReLS & 1.0513 & 0.0257\\
		\hline
		LMA & 1.2569 & 0.1132\\
		\hline
		LMAT & 1.1537 & 0.0818\\
		\hline
		IPR & 4.1943 & 36.2011\\
		\hline
		SIPR & -14.1352 & $<$ 1e-04\\
		\hline
		LL & 1.1878 & 0.0006\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.4700 & 0.0635\\
		\hline
	\end{tabular}
	\caption{Linear Data, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.7691 & 0.4216\\
		\hline
		ReLS & 1.7691 & 0.2789\\
		\hline
		LMA & -0.0161 & 0.1099\\
		\hline
		LMAT & 1.3810 & 8.5140\\
		\hline
		IPR & 5.8926 & 17.3418\\
		\hline
		SIPR & 14.9726 & $<$ 1e-04\\
		\hline
		LL & 1.9553 & 0.0164\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 0.2586 & 0.0266\\
		\hline
	\end{tabular}
	\caption{4th Order Model, $\rr$ Estimates and Standard Errors}
\end{table}

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r}
		\hline
		Model & Estimate & Std. Dev\\
		\hline
		LS & 1.2740 & 0.0971\\
		\hline
		ReLS & 1.2738 & 0.0334\\
		\hline
		LMA & 0.9503 & 0.1067\\
		\hline
		LMAT & 1.1959 & 0.0247\\
		\hline
		IPR & 1.2085 & 1.0031\\
		\hline
		SIPR & 1.3045 & 0.2300\\
		\hline
		LL & 1.2585 & 0.0010\\
		\hline
		MC & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.2588 & 0.0587\\
		\hline
	\end{tabular}
	\caption{Linear SIR Data, $\rr$ Estimates and Standard Errors}
\end{table}

\section{Variance Estimation Methods}\label{sec:sim-var-res}
We examine the variance methods used to estimate CIs for our estimates of $\rr$.  From the above results, we saw that methods such as LS, ReLS, LL, and MC result in very small CIs.  On the other hand, methods such as IPR have consistently large CIs, to the point of being uninformative.  The other methods CI width seems to depend heavily on the initial parameters used.  Unsurprisingly, an $\sigma_X$, $\sigma_Y$, $T$, or $N$ seems to be positively associated with a larger CI.  Having a larger percentage of initially infected individuals also seems to be associated with a larger CI.  We also see that having a very large $\rr$ value seems to have less effect on creating a large CI than does having a small value of $\rr$.  Perhaps surprisingly, the width of the CI seems to be independent of the error structure.

We estimate the variance through the delta method, block bootstrap, and posterior and compare it to the empirical variance of $\rr$ for each of the methods, using 1000 data sets generated from the baseline data structure.

\subsection{Repeated Sampling}

Table \ref{tab:rep-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 repetitions of the baseline conditions, with new noise generated each time. We want to compare the standard deviations of the estimates to the standard errors returned by each of the models for a single data set to asses how reliable those standard errors are. Overall, it appears than most of the models have standard error estimates that are larger than the actual standard deviation of the estimates under repeated sampling. In some cases, such as LL, LMA, and LMAT, the standard deviation and standard error estimates are fairly close. The least squares methods, LS and ReLS, on the other hand, tend to overestimate the error by an order of magnitude. Finally, SIPR and SB, have individual standard error estimates that are smaller than the standard deviation of repeated estimates, meaning that they understate how much error is in the estimates.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r|r|r}
		\hline
		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
		\hline
		LS & 2.0000 & 0.0003 & 1.9999 & 0.0058\\
		\hline
		ReLS & 2.0000 & 0.0003 & 1.9997 & 0.0052\\
		\hline
		LMA & 1.8497 & 0.3339 & 1.9974 & 0.3900\\
		\hline
		LMAT & 1.4562 & 0.3577 & 1.3980 & 0.6028 \\
		\hline
		IPR & 4.5214 & 0.3705 & 2.5930 & 9.0408\\
		\hline
		SIPR & 1.7973 & 0.0951 & 0.9251 & $<$ 1e-04 \\
		\hline
		LL & 2.0000 & 0.0003 & 1.9998 & 0.0002\\
		\hline
		MC & 0.9486 & $<$ 1e-04 & 0.9486 & $<$ 1e-04\\
		\hline
		SB & 1.1330 & 0.4163 & 1.0000 & 0.0523\\
		\hline
	\end{tabular}
        \caption{Mean $\rr$ Estimates and Std. Devs, Repeated Sampling}
        \label{tab:rep-samp}
\end{table}

\subsection{Delta Method}

\subsection{Block Bootstrap}

Table \ref{tab:bb-samp} shows the mean and standard deviation of $\rr$ estimates produced by the models under 100 block bootstrap repetitions of the baseline data. The data were first detrended using the predictions of the compartment sizes at each time point from each of the models on the original data set. The block bootstrap algorithm was then applied on to the residuals from the predictions using a fixed block length. Only 5 of the models provide predictions of the size of the compartments and thus allow for detrending of the data. The least squares and reparameterized model have a fairly accurate estimates of $\rr$ as well as very small standard errors. On the other hand, the linear approximation and smoothed IPR models both have very inaccurate estimates as well as very large standard errors of $\rr$.

\begin{table}[H]

	\centering
	\begin{tabular}[t]{l|r|r|r|r}
		\hline
		Model & Mean Estimate & Std. Dev. & Baseline Estimate & Baseline SE\\
		\hline
		LS & 1.9999 & 0.0003 & 1.9999 & 0.0058\\
		\hline
		ReLS & 1.9998 & 0.0003 & 1.9997 & 0.0052\\
		\hline
		LMA & -0.9975 & 21.7161 & 1.9974 & 0.3900\\
		\hline
		LMAT & 1.5138 & 1.0485 & 1.3980 & 0.6028 \\
		\hline
		SIPR & 0.7762 & 2.6624 & 0.9252 & $<$ 1e-04 \\
		\hline
	\end{tabular}
	\caption{Mean $\rr$ Estimates and Std. Devs, Block Bootstrap}
	\label{tab:bb-samp}
\end{table}

Figure \ref{fig:coverage} shows the proportion of repetitions for both repeated sampling and block bootstrapping in which the 95\% confidence intervals of the estimates produced by the models covers the true $\rr$. The least squares, reparameterized, and linear model approximation with all time points contain the true $\rr$ value in all of the confidence intervals. On the other hand, the smoothed IPR, Markov chain, and sequential Bayes models never cover the true $\rr$ in any of the confidence intervals.

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{images/coverage.jpeg}
		\caption{Proportion of coverage of true $\rr$ of confidence intervals produced by repeated sampling and block bootstrapping.}
		\label{fig:coverage}	
	\end{center}
      \end{figure}

      \subsection{Posterior}

      \section{Application to USA H1N1 data}\label{sec:real-data}
      We apply our methods using data from the 2009 H1N1 pandemic in the United States.

      \subsection{USA H1N1 Influenza 2009 Data}
      The data is from the Center of Disease Control and Prevention's (CDC) FluView \citep{cdc-fluview}.  The data source is ILINet and is the national level data.  We use the same date range as in \cite{towers2009}, namely Epiweeks 21-33 (May 23, 2009-August 22, 2009).  From the United States Census Bureau, we use the population estimate of the USA as of April 1, 2010 of $N=308,740,000$ individuals \citep{census-2010}.  The main features in this data set are weighted influenza like illness (wILI) for a given week (Epiweek).  The wILI values are estimated based on the number of patients diagnosed with influenza like illness and weighted based on the reporting locations.  To obtain weekly incidence from wILI, we use $J(t) = \textnormal{wILI}(t) / 100 \cdot N$.  Then to obtain prevalence $I(t)$, we use the relation of $I(t) = \gamma^{-1}J(t)$, where $\gamma$ is the inverse of the expected time to recovery, which we here use as $\gamma^{-1} = \frac{3}{7}=.429$ weeks \citep{vespignani2007}.  Then $S(t) = N - \sum_{s=0}^{t}J(s)$, the total population size minus the cumulative sum of the incidence, and $R(t) = N - S(t) - I(t)$.  As prior knowledge of $\gamma$ is used to estimate our data for $I$, we expect it to influence our estimates of $\rr$.  The data is summarized in Table \ref{tab:h1n1-data}.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[H]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
Feature       & Notes                                                                   \\ \midrule
Disease       & H1N1 Pandemic Influenza                                                 \\
Source        & CDC FluView (\url{https://www.cdc.gov/flu/weekly/})      \\
Year          & 2009                                                                    \\
Date(s)          & May 23-August 22 (Epiweeks 21-33)                                       \\
Location      & USA                                                                     \\
Population    & $N$=308,740,000 people                                                  \\
  Avg. Duration $\left ( \gamma^{-1}\right )$ & $\frac{3}{7}$ weeks\\
  Raw data & Weekly wILI reports\\
  Final data & S$(t)$, I$(t)$, R$(t)$ \\\bottomrule
\end{tabular}
\caption{Summary of the data used to estimate $\rr$ during the 2009 H1N1 influenza pandemic in the USA.  We follow the data collection procedure described in \cite{towers2009}.}
\label{tab:h1n1-data}
\end{table}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{images/h1n1-sir.pdf}
  \caption{Data from H1N1 influenza 2009 in the USA.  Weekly wILI is reported, which we transform into S, I, and R counts for epiweeks 21-33 using $\gamma^{-1} = 3/7$ weeks.}
  \end{figure}

      \subsection{Results}
      The results are shown below.  \cite{towers2009} report an estimate of $\hat{\rr} \in [1.50, 1.70]$.

\section{Discussion}\label{sec:discussion}

%% say what we said
%% say it
%% say it again

%% what is r0 and why it is important
In this paper, we examine $\rr$, the average number of secondary infections a primary individual generates when introduced into a completely susceptible population, in the context of the SIR model.  The quantity $\rr$ summarizes an epidemic into a single estimate in the hopes that infectious disease outbreaks may be compared to one another.  A value of $\rr > 1$ indicates an outbreak of a disease whereas a value of $\rr < 1$ indicates that no outbreak will occur.  Thus, it is critical to not only estimate the value of $\rr$ but also a corresponding CI.  Overestimating $\rr > 1$ during an ongoing epidemic may result in a waste of resources in prevention and control measures; underestimating $\rr < 1$ may result in a worldwide crisis.

In Table \ref{tab:r0-real-ex}, we provide estimates for $\rr$ that have appeared over the past three decades for HIV, Zika, Ebola, influenza, H1N1 influenza, and measles in different locations throughout the world.  Each of these estimates is greater than $1$, which makes sense as we tend to care more when a disease has the potential outbreak.  The largest $\rr$ point estimate we have seen is for measles at $\rr=17$ in England in the 20th century, and the smallest is $\rr=1.08$ for HIV in Portugal in the 21st century.
%% r0 as a property of the model
The quantity $\rr$ may summarize an epidemic, accounting for temperature, seasonality, location, and genetics of a disease, but it is still a property of the model.  Namely, how individuals progress through an epidemic (e.g. susceptible to infectious to recovered compartments) influences how we view $\rr$.  We thus should compare estimates for $\rr$ for a fixed epidemic model.  Here, we limit our analysis to estimates of $\rr$ derived from the SIR model.  Then $\rr = \frac{\beta}{\gamma}$ where $\beta$ is the rate of infection and $\gamma$ is the rate of recovery.  As a consequence, we need to estimate $\rr$ from not only the number of infections but also the number of individuals recovering at each time step.

%% the methods we chose and why we chose them
Additionally, the SIR model assumes monotoncity of S and R compartments which can make estimating $\rr$ more difficult, especially in terms of the CIs since the usual assumption of Gaussian noise may be violated due to the non-symmetric nature of the noise.  We examine \wxxsir methods to estimate $\rr$ which include least squares optimization, polynomial approximations, estimates from incidence and variance, a log-linear model, a Markov chain method, and a simple Bayesian method.  Although, this is not an exhaustive list of methods used to estimate $\rr$ from the SIR model, we believe it covers common and popular methods used throughout the decades.  We overview the details of methods in Section \ref{sec:methods}.

%% the importance of variance and testing
Each of the \wxxsir methods results in an estimate of $\rr$, $\hat{\rr}$.  We also estimate a 95\% CI for the  $\rr$ estimate.  Although, many researchers report intervals via sensitivity analysis (how the estimate changes with small variations in the underlying parameters), the ranges of these parameters often do not incorporate any distributional assumptions about these estimated ranges and hence do not correspond to 95\% CIs.  We provide three statistical methods to estimate the standard error of $\hat{\rr}$ and use the interval $\left [\hat{\rr} \pm 2\hat{SE}\left (\hat{\rr} \right ) \right ]$ as our 95\% CI.  We hope that these intervals will provide guidance on how to create more meaningful intervals when estimating $\rr$ and discerning which diseases are more severe than others, and whether the true $\rr >1$.

We compare the the estimates of $\rr$ for the \wxxsir methods using a series of simulations.  We generate different simulated data sets under different noise assumptions and estimate $\rr$ using each of these data sets.  We then analyze estimates of $\rr$ while varying individual parameters of the model including the infection and recovery parameters, the total amount of data, the initial percent of infectious and susceptible individuals, the standard deviation of the noise used to generate the simulated data, and the total number of individuals in a population.  We also analyze the model under misspecification, that is when a model other than the SIR model is used to generate the simulated data sets.  Finally, we compare our variance methods to one another and analyze the coverage of our methods.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%RESULTS
%%%%%%%%%%%%%%%%%

Overall, we recommend based on our analysis:
\begin{itemize}
  \item LS and ReLS provide the most accurate estimates of $\rr$
  \item LL results in close point estimates but occasionally has too small standard errors
  \item The specific MC and SB methods presented here do not seem to fit our generated data well
  \item There is not much difference in estimations of $\rr$ due to different error generating processes
  \item Very small values of $\gamma$ can result in very poor estimates of $\rr$
  \item 100 data points seems to be a reasonable number of data points to have enough power for testing that $\rr > 1$ and 50 seems to be too few
  \item The larger the number of initially infectious individuals is, the smaller the CIs for $\rr$ are
  \item The noisier the epidemic data is, the larger our CIs for $\rr$ are
  \item Estimations of $\rr$ are fairly robust to the misspecified data sets we presented, at least for the methods of LS, ReLS, and LL
  \item Block bootstrap matches the empirical variance through repeated sampling.
  \end{itemize}

$\rr$ is a difficult quantity to estimate, but we hope that our methods of estimation and simulations provide some guidance for researchers on how to generate statistically reliable estimates for $\rr$ and its CIs.  Our analysis is limited in scope to the SIR model, and in the future we would like to see how methods of estimating $\rr$ compare for models such as the SEIR and beyond.

%% results of simulations:
%% - data
%% - what parameters make a difference
%% - what methods perform best when
%% - misspecification
%% variance
%% beyond the sir model

The SIR estimation methods generally give reliable estimates when the data are from an underlying SIR model with little or no noise. However, the estimates from the various methods are generally much worse when the magnitude of the noise is larger, when the parameters of the model are unusual, and when the model is misspecified. We also see a large difference between the $\rr$ estimates and standard errors on the exact same data sets as well, even though all derived from the same model. Thus, when estimating $\rr$ based on an SIR or any other kind of compartment model, we need to take careful consideration.

\bibliographystyle{apa}%Choose a bibliograhpic style
\bibliography{Master}

\appendix

\section{Beyond the SIR Model}

\subsection{Exponential Growth (EG)}\label{sec:expgrowth}
%% nishiura, majumder, IDEA model -- sum of squares (santillana2016)
%% An IDEA for short term outbreak projection: nearcasting using the basic reproduction number.  (fisman2013, fisman2014)
%% chowell 2008 - parametric bootstrap for a model they made up, but sort of empirical.
Many researchers employ an assumption of exponential growth of the number infectious through times $t=0, 1, \dots, T^* <T$ \citep{wallinga2007generation,fisman2014,nishiura2016,majumder2016,towers2016}.
\cite{wallinga2007generation} report that the effective reproduction number $\mathcal{R}_t$ and hence the initial reproduction number $\rr$ may derived using the fact that infection ``counts increase exponentially in the initial phase of an epidemic.''  We then have to estimate $r$, the \textit{per capita} change in the number of new cases per unit of time and $\omega$ the serial interval, the distribution of time between a primary and secondary infection. Then, we have
\begin{align}\label{eq:lotka}
\rr = \exp{(r \omega)}.
\end{align}
Equation \eqref{eq:lotka} is derived from a demographic view using the Lotka-Euler survival equations which come from the fields of demography, ecology, and evolutionary biology.

\cite{wallinga2007generation}, instead, use a moment generating function expression for $\rr$.  With $\omega(t)$ as the serial interval, then
\begin{align*}
\rr^{-1} &= \frac{1}{M(-r)} = \int_{a=0}^\infty e^{-rt}\omega(t)dt.
\end{align*}
\cite{wallinga2007generation} specifically derive an estimate for $\rr$ for the SIR model.  They assume that the rate of leaving the infectious stsate, $\gamma$ is constant as is the rate of making contacts during the infectious stage.  Then, the duration of a serial interval is an exponential distribution with mean $\gamma^{-1}$, the moment generating function is then,
\begin{align*}
  M(-r) = \frac{\gamma}{\gamma + r},
\end{align*}
and the resulting estimate of $\rr$ is
\begin{align}\label{eq:r0-eg}
  \hat{\rr} = 1 + \frac{\hat{r}}{\hat{\gamma}},
\end{align}
where $\hat{r}$ and $\hat{\gamma}$ are estimates of the $r$ and $\gamma$, respectively.  The parameter $\hat{\gamma}$ is often assumed to be fixed or tried with a sequence of serial intervals \citep{majumder2016}.  We may estimate $\hat{r}$ using ordinary linear regression, namely,
\begin{align*}
  \log \left  (\frac{Y(t)}{Y(0)} \right )&= rt + \epsilon_t\\
  \epsilon_t &\sim N(0, \sigma^2),
\end{align*}
which results in
\begin{align}\label{eq:exp-growth}
  \hat{r} &= \frac{\sum_{t=0}^{T^*}\left  (\frac{Y(t)}{Y(0)} \right )}{\sum_{t=0}^{T^*}t}.
\end{align}

For the above estimate of $\rr$, we observe the duration of serial intervals in a period of exponential growth of the number of infectious individuals.  Deciding when exponential growth occurs and which data points to use to estimate $\rr$ may be difficult.  We can use sensitivity analysis to see how the estimate $\rr$ changes as we increase or decrease $T^{*}$, the number of time points used. The reader may notice some similarities to the estimate for $\rr$ in Equation \eqref{eq:r0-eg} and from the log-linear model in Equation \eqref{eq:harko_lin}.  This method is a simplification of the SIR model as the $R$ compartment is not used to estimate $r$, only the recovery rate $\gamma$.

An advantage to this method are its (and similar adaptations) prevalence in the literature, which allows one to to compare estimates $\rr$ directly to one another.  Another advantage is its flexibility.  Here, we assumed the serial interval has an exponential distribution, but this is not a requirement and other distributions may be used (see \cite{wallinga2007generation}).  However, a disadvantage to this method is that is a simplification of the SIR model and we have to decide how many time points to use to estimate the exponential growth rate of infectious.  Another disadvantage to this method is that one also has to decide how to treat the additional paramter $\gamma$, whether to treat it as fixed, a range of fixed values, or even having a probability distribution.

\subsection{Incidence Decay and Exponential Adjustment (IDEA)}\label{sec:idea}
The Incidence Decay and Exponential Adjustment (IDEA) model is a way to account for a declining effective reproduction number of time due to both depletion of susceptible individuals and spontaneous and planned control activities and behaviors'' \citep{fisman2013}.  They model the incidence counts during the $t$th interval, $J(t),$ as
\begin{align*}
  J(t) &= \left [ \frac{\rr}{(1 + d)^t}\right ]^t,
\end{align*}
where $d$ is a dampening or discount factor.  As $d$ increases, the incidence at time $t$  decreases. The estimate for $\rr$ is obtained by minimizing the sum of squares over $\rr$ and $d$,
\begin{align*}
  \hat{\rr} &=  \arg \min_{\rr,d} \sum_{t=0}^{T^*}\left (J_{obs}(t) - J(t) \right)^2.
\end{align*}
The method presented in \cite{fisman2013} is a mathematical, rather than statistical model.  Sensitivity analysis is conducted with respect to $T^*$, $d$, and $\rr$, and lower and upper bounds are reported as a result of this analysis.

Advantages of this method include its usefulness in estimating $\rr$ \citep{fisman2014,majumder2016}.  Additionally, this model may be employed to address the effect of control measures and better estimate the final size of an epidemic.  A disadvantage to this model is that its upper and lower bounds do not correspond to CIs.  Also, in this method we must select the interval size of the incidence counts.

\end{document}

%%%%%%%%%%%%%%%%%%%%%%
\textbf{Title:}

\textbf{Author:}

\textbf{Citation:}

\textbf{Major themes:}

\textbf{Notes:}
\\
%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\begin{center}
\includegraphics[width=4in]{images/mvhw7_3c.pdf}
\end{center}
\caption{
Biplots of the different continuous variables and their correlations.}\label{fig3c}
\end{figure}

%% two pictures whoa

\begin{figure}[h]
\centering

\begin{figure}[h]
\centering

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
  \caption{Histograms of Arrival Delay and continuous covariates.  Arrival delay seems to have a right skewed distribution.  This may indicate that we will be transforming this variable later on.  After transforming Air Time and Distance by a log transformation, we don't really seem to have many outliers in our covariates.  We seem to have outliers in the CRS Dep. Time and Arrival Time; however, time is cyclical and so these are not, in fact outliers.}
  \label{hists}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{images/mt_eda_cont_hists.pdf}
  \caption{\textcolor{red}{Placeholder}}
  \label{tabs1}
\end{subfigure}
\caption{}
\end{figure}

\begin{table}
\begin{tabular}{l | a | b | a | b}
\hline
\rowcolor{LightCyan}
\mc{1}{}  & \mc{1}{x} & \mc{1}{y} & \mc{1}{w} & \mc{1}{z} \    \hline
variable 1 & a & b & c & d \    variable 2 & a & b & c & d \\ \hline
\end{tabular}
\end{table}

%%
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{images/resids_full.pdf}
  \caption{}
  \label{residsf}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{images/diags_full.pdf}
  \caption{ }
  \label{diagsf}
\end{subfigure}
\caption{}
\end{figure}